---
title: 'ETC5523: Communicating with Data'
subtitle: "Workshop 8: Build and document `pokemonvgc`"
author: "Michael Lydeamore"
date: "Week 8"
format:
  unilur-html:
    output-file: index.html
    embed-resources: true
  unilur-html+solution:
    output-file: solution.html
    embed-resources: true
---

## 🎯 Objectives

By the end of this workshop, you will be able to:

* distinguish raw data inputs from data shipped in an R package;
* document and export a package function with `roxygen2`;
* replace repeated `package::function()` calls with selective imports;
* write a README quick start; and
* check, install and use a package.

## Get the starter package

Clone <https://github.com/MikeLydeamore/pokemonvgc> and open
`pokemonvgc.Rproj` in Positron or RStudio.

You can clone through the **Git: Clone** command, or run this in a terminal:

```bash
git clone https://github.com/MikeLydeamore/pokemonvgc.git
```

Install the workshop tools if you do not already have them:

```{r}
#| eval: false

install.packages(c(
  "devtools", "dplyr", "remotes", "roxygen2", "rmarkdown", "usethis"
))
```

::: {.callout-important}
Run the remaining commands with `pokemonvgc` open as your current project.
Do not create another package inside it.
:::

## Exercise 8A: Tour the package (5 minutes)

Find each of these components and write one sentence describing its role:

| Component | Question to answer |
|---|---|
| `R/` | What code will be available to package users? |
| `data/` | Which objects will be loaded with the package? |
| `data-raw/` | Which inputs and scripts produce those objects? |
| `DESCRIPTION` | What does this tell R and a potential user? |
| `README.Rmd` | Which reader encounters this document first? |

Load the package without installing it:

```{r}
#| eval: false

devtools::load_all()
```

Use `pokemon_elo` and `recent_champions` to answer:

1. How many rating observations and distinct players are included?
2. Which recent champion has no rating history in this snapshot?
3. Can you open `?recent_players` yet? Why or why not?

::: {.unilur-solution}

```{r}
#| eval: false

nrow(pokemon_elo)
dplyr::n_distinct(pokemon_elo$player_id)

recent_champions |>
  dplyr::anti_join(
    dplyr::distinct(pokemon_elo, player_id),
    by = "player_id"
  )
```

There are 2,043 observations for 51 players. Takuma Yamazaki is in the
champion metadata but not the rating snapshot. The function has no help page
because its roxygen block has not been written and generated yet.

:::

## Exercise 8B: Rebuild the package data (7 minutes)

Open `data-raw/pokemon-elo.R` and follow the data from its prepared inputs to
the objects saved by `usethis::use_data()`.

Then run the script from the package root:

```{r}
#| eval: false

source("data-raw/pokemon-elo.R")
```

The calculation may take several seconds. Confirm that it recreates:

* `data/pokemon_elo.rda`; and
* `data/recent_champions.rda`.

Discuss with a partner:

* Why is the match history kept in `data-raw/` rather than shipped to users?
* Why is `recent_champions` stored as package data rather than embedded in
  `recent_players()`?
* Which file would you edit if the champion lookup changed?

::: {.unilur-solution}

The match history is an input to the reproducible build process; users need
the smaller derived rating history, not all source records. Keeping the
champion lookup as data separates metadata from the function's logic and makes
it available for other analyses. Update `data-raw/pokemon-elo.R`, rerun it,
and commit the regenerated `.rda` file rather than editing `data/` directly.

:::

## Exercise 8C: Document and export the function (10 minutes)

Open `R/recent-players.R`. Read the function as evidence before writing its
documentation. What does one input row represent? When is a player eligible?
What does one output row represent?

Add a `roxygen2` block containing:

* a task-oriented title and short description;
* `@param` entries for `ratings` and `min_events`;
* an `@return` entry stating what one row represents;
* two runnable `@examples`, one using a non-default threshold; and
* `@export`.

Generate the documentation and reload the package:

```{r}
#| eval: false

devtools::document()
devtools::load_all()
?recent_players
```

Inspect the new files in `man/` and `NAMESPACE`. Do not edit either generated
file by hand.

::: {.unilur-solution}

One possible first version is:

```r
#' Find each eligible player's most recent rating
#'
#' Keeps the latest rating for every player with at least `min_events`
#' observations in the supplied rating histories.
#'
#' @param ratings A data frame containing `player_id`, `player_name`, `date`,
#'   and `rating`.
#' @param min_events Minimum number of observations a player must have.
#' @return A data frame with one row per eligible player, including their most
#'   recent date and rating and their total number of observations.
#' @examples
#' recent_players(pokemon_elo)
#' recent_players(pokemon_elo, min_events = 10)
#' @export
```

After `devtools::document()`, `NAMESPACE` should contain
`export(recent_players)` and `man/recent_players.Rd` should exist.

:::

## Exercise 8D: Use selective imports (8 minutes)

The first version of the function uses `dplyr::` repeatedly. `dplyr` is
already recorded under `Imports` in `DESCRIPTION`. Now use roxygen to name
only the functions imported into the package namespace.

1. Add this line to the roxygen block:

   ```r
   #' @importFrom dplyr filter group_by mutate n slice_max ungroup
   ```

2. Remove the six `dplyr::` prefixes from the function body.
3. Run `devtools::document()` again.
4. Inspect the generated `NAMESPACE`, then run `devtools::load_all()`.

Explain the different jobs performed by:

* `Imports: dplyr` in `DESCRIPTION`;
* `@importFrom dplyr ...` in the roxygen block; and
* `importFrom(dplyr, ...)` in `NAMESPACE`.

::: {.unilur-solution}

The finished function can read:

```r
#' @importFrom dplyr filter group_by mutate n slice_max ungroup
#' @export
recent_players <- function(ratings, min_events = 15) {
  ratings |>
    group_by(player_id, player_name) |>
    mutate(events = n()) |>
    slice_max(date, n = 1, with_ties = FALSE) |>
    ungroup() |>
    filter(events >= min_events)
}
```

`DESCRIPTION` records that the package depends on `dplyr`. The roxygen tag is
the editable source declaring which names the code uses. `document()` turns
that tag into machine-readable `importFrom()` entries in `NAMESPACE`.

:::

## Exercise 8E: Test from a user's perspective (5 minutes)

Run the documented function and identify the current top three eligible
players:

```{r}
#| eval: false

current_ratings <- recent_players(pokemon_elo)

current_ratings |>
  dplyr::slice_max(rating, n = 3, with_ties = FALSE)
```

Check that:

* the result has one row per player;
* every returned player has at least 15 observations; and
* changing `min_events` changes the eligibility rule.

Ask a partner to use only `?recent_players` to explain the `events` column and
the result of changing `min_events`. Revise the first unclear sentence.

::: {.unilur-solution}

The top three are Wolfe Glick, Eric Rios and Paul Chua. With the default
threshold, the function returns 51 rows. Here `events` counts rating
observations in the supplied data, so the documentation should not imply that
it counts every event a player has ever attended.

:::

## Exercise 8F: Write the README quick start (7 minutes)

Complete the TODOs in `README.Rmd`. A first-time reader should learn:

1. what the package helps them do;
2. how to install it from GitHub;
3. how to find the latest eligible ratings; and
4. why the Elo values require care when interpreted.

Use this copyable installation command:

```r
remotes::install_github("MikeLydeamore/pokemonvgc")
```

Keep the quick start short and runnable. Then regenerate `README.md`:

```{r}
#| eval: false

devtools::build_readme()
```

::: {.unilur-solution}

A useful quick start might include:

```r
library(pokemonvgc)

recent_players(pokemon_elo) |>
  dplyr::slice_max(rating, n = 3, with_ties = FALSE)
```

The surrounding text should say that this returns the three highest latest
ratings among players with at least 15 observations. It should also describe
the ratings as estimates from an incomplete match archive, not an official
ranking.

:::

## Exercise 8G: Check and install (8 minutes)

Check the package as a complete product:

```{r}
#| eval: false

devtools::document()
devtools::check()
```

Fix every error and warning. Read every note and decide whether it needs
action. When the check passes, install the package:

```{r}
#| eval: false

devtools::install()
```

Restart R, then confirm that the installed package works without
`devtools::load_all()`:

```{r}
#| eval: false

library(pokemonvgc)
?recent_players
recent_players(pokemon_elo)
```

## Completion check

You have finished when:

* the data-raw script reproduces both packaged datasets;
* `recent_players()` has a useful help page and is exported;
* `DESCRIPTION` and generated `NAMESPACE` record the selective imports;
* `README.md` contains a copyable installation command and runnable example;
* `devtools::check()` reports no errors or warnings; and
* the installed package works in a clean R session.
