---
title: 'ETC5523: Communicating with Data'
subtitle: "Workshop 8: Documentation as an invitation"
author: "Michael Lydeamore"
date: "Week 8"
format:
  unilur-html:
    output-file: index.html
    embed-resources: true
  unilur-html+solution:
    output-file: solution.html
    embed-resources: true
---

## 🎯 Objectives

By the end of this workshop, you will be able to:

* describe a function from the user's perspective;
* write useful reference documentation and examples; and
* create a short README quick start that helps someone act.

## 1. Quickwrite: a cold start (5 minutes)

Recall a time when instructions assumed something you did not know. What did the writer assume? What single addition would have helped you begin?

## 2. Define the user and their task (5 minutes)

You are documenting this function for an R user who understands vectors but has not read its source code:

```{r}
#| eval: false
#| echo: true

percent_change <- function(old, new, digits = 1) {
  if (any(old == 0)) {
    stop("old cannot contain zero")
  }
  round((new - old) / old * 100, digits = digits)
}
```

Before documenting individual arguments, write:

* the task the user is trying to complete;
* what the function returns; and
* one situation in which the function should not be used without care.

## 3. Read the function as evidence (5 minutes)

Run these examples mentally and predict the output or error:

```{r}
#| eval: false
#| echo: true

percent_change(80, 100)
percent_change(c(80, 50), c(100, 45), digits = 0)
percent_change(0, 20)
```

List anything a user could misunderstand: direction, units, vector behaviour, rounding or the zero restriction.

## 4. Draft the reference documentation (12 minutes)

Write a `roxygen2` block containing:

* a specific title;
* a one- or two-sentence description;
* `@param` entries for `old`, `new` and `digits`;
* an `@return` entry that states the type and meaning;
* two `@examples`, including a vector example; and
* `@export`.

Aim to answer user questions, not restate the argument names.

## 5. Write the README quick start (8 minutes)

Write a short section with the heading **Calculate a percentage change**. In no more than 100 words:

1. state why someone would use the function;
2. show one runnable call;
3. show or explain the result; and
4. mention the most important restriction.

## 6. Cold-read test (7 minutes)

Exchange the documentation, but not the original function, with a partner. The reader answers:

* Which value is treated as the baseline?
* What does a negative result mean?
* Can the function accept vectors?
* What happens when the baseline is zero?

Revise the first place where the documentation failed to answer clearly.

## 7. Writer's note (3 minutes)

Complete: "Good documentation begins with the user's ______, not the code's ______."

::: {.unilur-solution}
## Instructor notes and example core description

A useful opening is:

> Calculate the percentage change from an old value to a new value. Positive results indicate an increase relative to `old`; negative results indicate a decrease.

The return documentation should identify a numeric vector measured in percentage points of change, rounded to `digits`. Encourage examples that teach interpretation, not merely prove the function runs.
:::
