---
pagetitle: "ETC5523: Communicating with Data"
subtitle: "CSS, AI-assisted coding, and polished ggplots"
author: "Michael Lydeamore"
email: "michael.lydeamore@monash.edu"
date: "Week 7"
department: "Department of Econometrics and Business Statistics"
unit-url: "cwd.numbat.space"
footer: "ETC5523 Week 7"
format:
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    multiplex: false
    theme: ../assets/monash.scss
    show-slide-number: all
    show-notes: false
    slide-number: c
    controls: true
    width: 1280
    height: 720
    auto-stretch: false
    code-line-numbers: false
    css: [../assets/tachyons-addon.css, ../assets/custom.css, week7.css]
    include-after-body: "../assets/after-body.html"
    chalkboard:
      boardmarker-width: 5
      buttons: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(here)
library(ggrepel)
library(scales)

knitr::opts_chunk$set(
  fig.path = "images/index/",
  fig.width = 8,
  fig.height = 4.5,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 3,
  dev = "svg",
  dev.args = list(bg = "transparent"),
  warning = FALSE,
  message = FALSE
)

brand <- c(ink = "#222222", muted = "#B8B8B8", accent = "#006DAE")

birthplace <- read_csv(here("data/census-birthplace.csv"), show_col_types = FALSE) |>
  filter(
    census == 2021,
    !birth %in% c("Total", "Australia", "Other", "Not Stated")
  ) |>
  slice_max(count, n = 7) |>
  mutate(
    birth = fct_reorder(birth, percentage),
    focus = if_else(birth == "India", "Focus", "Context")
  )

housing <- ggplot2::txhousing |>
  filter(
    city %in% c("Austin", "Dallas", "Houston", "San Antonio"),
    year >= 2005,
    !is.na(median)
  ) |>
  mutate(
    date = as.Date(sprintf("%d-%02d-01", year, month)),
    focus = if_else(city == "Austin", "Focus", "Context")
  )

housing_end <- housing |>
  group_by(city) |>
  slice_max(date, n = 1, with_ties = FALSE) |>
  ungroup()

airport_year <- read_csv(
  here("data/melb-airport-max-temp-data.csv"),
  show_col_types = FALSE
) |>
  transmute(
    year = Year,
    max_temp = `Maximum temperature (Degree C)`
  ) |>
  group_by(year) |>
  summarise(mean_max = mean(max_temp, na.rm = TRUE), .groups = "drop") |>
  filter(is.finite(mean_max))

airport_reference <- mean(airport_year$mean_max)
airport_high <- airport_year |>
  slice_max(mean_max, n = 1, with_ties = FALSE)

theme_cwd <- function(base_size = 16) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title.position = "plot",
      plot.caption.position = "plot",
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      axis.title = element_text(colour = brand[["ink"]]),
      axis.text = element_text(colour = brand[["ink"]]),
      plot.title = element_text(face = "bold", colour = brand[["ink"]]),
      plot.subtitle = element_text(colour = "grey35"),
      plot.caption = element_text(colour = "grey40", hjust = 0),
      legend.position = "bottom",
      plot.margin = margin(10, 36, 10, 10)
    )
}
```

## <br>[`r rmarkdown::metadata$pagetitle`]{.monash-blue} {#etc5523-title background-image="images/bg-01.png"}

### `r rmarkdown::metadata$subtitle`

Lecturer: *`r rmarkdown::metadata$author`*

`r rmarkdown::metadata$department`

::: tl
<br>

<ul class="fa-ul">
<li>[<i class="fas fa-envelope"></i>]{.fa-li}`r rmarkdown::metadata$email`</li>
<li>[<i class="fas fa-calendar-alt"></i>]{.fa-li} `r rmarkdown::metadata$date`</li>
<li>[<i class="fa-solid fa-globe"></i>]{.fa-li}<a href="`r rmarkdown::metadata[["unit-url"]]`">`r rmarkdown::metadata[["unit-url"]]`</a></li>
</ul>
:::

## {#aim background="white"}

::: {.callout-important}

## Aim

By the end of today, you should be able to:

* apply small, scoped CSS changes to Quarto HTML output;
* brief an LLM to produce useful CSS or `ggplot2` code;
* inspect and verify generated code rather than accepting it on trust; and
* implement reusable techniques that turn a correct plot into a purposeful one.

:::

## The through-line

::: {.r-fit-text}

**Visual intention** → **technical vocabulary** → **code** → **inspection** → **revision**

:::

. . .

An LLM can accelerate the middle of this process. It cannot decide whether the result communicates honestly and clearly.

## Today extends Week 6

::: {.columns}
::: {.column width="48%"}

### Week 6: design judgement

* What is the message?
* Which comparison matters?
* Where should attention go?
* Is the result honest and accessible?

:::
::: {.column width="48%"}

### Week 7: implementation

* How do I target the right element?
* How do I encode emphasis in code?
* How do I label without clutter?
* How do I get useful help from an LLM?

:::
:::

## Part 1: CSS and AI-assisted coding {.center background-color="#006DAE"}

## HTML is structure; CSS is appearance

```html
<section class="finding">
  <h2>Reliable service matters most</h2>
  <p id="result">Reliability was selected by 46% of commuters.</p>
</section>
```

. . .

You only need three pieces of HTML vocabulary today:

| Kind | Example | CSS target |
|---|---|---|
| element | `<h2>` | `h2` |
| class | `class="finding"` | `.finding` |
| identifier | `id="result"` | `#result` |

## A CSS rule answers two questions

```css
.finding {
  border-left: 8px solid #006DAE;
  padding-left: 1rem;
}
```

::: {.columns}
::: {.column width="48%"}

### What should change?

`.finding` is the **selector**.

:::
::: {.column width="48%"}

### How should it change?

Each **property: value** pair is a declaration.

:::
:::

## Scope selectors to the smallest useful target

```css
/* Every second-level heading */
h2 { color: #006DAE; }

/* Headings inside a finding */
.finding h2 { color: #006DAE; }

/* One unique result */
#result { font-weight: 700; }
```

::: {.callout-warning .fragment}

`*`, `div`, `p`, and `h2` can affect far more of a document than intended. Generated CSS often fails because its selector is too broad—or does not exist in the HTML at all.

:::

## The small set of properties you will use often

| Purpose | Properties |
|---|---|
| colour | `color`, `background-color`, `border-color` |
| type | `font-family`, `font-size`, `font-weight`, `line-height` |
| space | `margin`, `padding`, `gap` |
| edges | `border`, `border-radius` |
| alignment | `text-align`, `display`, `align-items` |

. . .

Start with the visual problem. Look up properties when you need them.

## Use colour as a named design decision

::: {.columns}
::: {.column width="50%"}

### CSS

```css
:root {
  --ink: #222222;
  --muted: #B8B8B8;
  --accent: #006DAE;
}

.finding {
  color: var(--ink);
  border-color: var(--accent);
}
```

:::
::: {.column width="50%"}

### R

```r
brand <- c(
  ink = "#222222",
  muted = "#B8B8B8",
  accent = "#006DAE"
)
```

:::
:::

The same visual vocabulary can travel across a webpage and its plots.

## Put reusable CSS in a file

```yaml
---
format:
  html:
    css: styles.css
---
```

. . .

Use an inline CSS block for a quick experiment:

````markdown
```{css}
.finding { border-left: 8px solid #006DAE; }
```
````

. . .

Move stable rules into `styles.css` so they can be reused and reviewed.

## Inspect the page before changing the code

1. Right-click the element and choose **Inspect**.
2. Identify its element, classes, and surrounding container.
3. Try a declaration in the browser's Styles panel.
4. Move the working rule into the source CSS.
5. Render again and check more than one page or screen size.

::: {.callout-tip .fragment}

Browser experiments are temporary. They help you discover the rule; the source file makes the change permanent.

:::

## “Make it look better” is not a technical brief

::: {.columns}
::: {.column width="48%"}

### Under-specified

> Make this callout look better with CSS.

. . .

Likely result: arbitrary colours, excessive decoration, and selectors that may not match the document.

:::
::: {.column width="48%"}

### Actionable

> In this Quarto HTML page, style only elements with class `.finding`. Add an 8 px blue left border and 1 rem of left padding. Do not change global headings or use `!important`. Return the smallest CSS rule and explain each declaration.

:::
:::

## A useful LLM brief has four parts

::: {.columns}
::: {.column width="24%"}

### Target

What should change?

:::
::: {.column width="24%"}

### Context

What code and output format matter?

:::
::: {.column width="24%"}

### Constraints

What must remain unchanged?

:::
::: {.column width="24%"}

### Check

How will you recognise success?

:::
:::

. . .

Request the **smallest change**, then ask what each selector, layer, or declaration does.

## Generated code can look plausible and still be wrong

::: {.columns}
::: {.column width='50%'}
### Model response

```css
.important-finding h2 {
  color: #006DAE !important;
}
```
:::
::: {.column width='50%'}

### Inspect it

* Does `.important-finding` exist in the supplied HTML?
* Is the `h2` the intended target?
* Why is `!important` necessary?
* Does the rule work after a clean render?
:::
:::



## A six-step workflow for AI-assisted coding

1. Supply the smallest relevant source and output context.
2. State the visual intention.
3. Specify constraints and an acceptance check.
4. Request a minimal edit and an explanation.
5. Render, inspect, and test the result.
6. Describe the observed problem in the next prompt.

::: {.callout-important .fragment}

Do not ask whether the code “should work”. Run it and collect evidence.

:::

## Live demonstration

Use the workflow to request one change:

> On a Quarto RevealJS slide, style only `.key-result` blocks with a Monash blue left border, comfortable internal spacing, and no background fill. Preserve the existing theme and return only the CSS plus a two-sentence explanation.

. . .

::: {.key-result}
**One selector, one job**

Only this block should change.
:::

. . .

Then verify:

* the class exists;
* the rule is scoped;
* the properties are valid;
* the slide still works at 1280 × 720.


## Part 2: A polished-plot cookbook {.center background-color="#006DAE"}

## Recipe 1: highlight one bar without depending on order

One category carries the message, but every bar currently receives equal attention.

```{r}
#| label: bar-before
#| echo: false
#| fig-height: 3.5
ggplot(birthplace, aes(percentage / 100, birth)) +
  geom_col(fill = brand[["muted"]], width = 0.72) +
  scale_x_continuous(labels = label_percent(accuracy = 0.1)) +
  labs(x = "Share of Australian residents", y = NULL) +
  theme_minimal(base_size = 16)
```

## Recipe 1: create an explicit focus variable

```r
plot_data <- plot_data |>
  mutate(focus = if_else(category == "India", "Focus", "Context"))

focus_colours <- c(
  "Context" = "#B8B8B8",
  "Focus" = "#006DAE"
)

ggplot(plot_data, aes(value, category, fill = focus)) +
  geom_col() +
  scale_fill_manual(values = focus_colours, guide = "none")
```

. . .

A **named** vector matches colours to meanings even if the rows or factor levels move.

## India ranked third among all countries of birth

```{r}
#| label: highlight-one-bar
#| echo: false
#| fig-height: 4.2
focus_colours <- c("Context" = brand[["muted"]], "Focus" = brand[["accent"]])

ggplot(birthplace, aes(percentage / 100, birth, fill = focus)) +
  geom_col(width = 0.72) +
  geom_text(
    aes(label = percent(percentage / 100, accuracy = 0.1)),
    hjust = -0.15, colour = brand[["ink"]], size = 4.5
  ) +
  scale_fill_manual(values = focus_colours, guide = "none") +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    expand = expansion(mult = c(0, 0.14))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    x = "Share of Australian residents", y = NULL,
    title = "India ranked third among all countries of birth",
    subtitle = "Australia ranked first and is omitted from this chart",
    caption = "Source: Australian Census 2021"
  ) +
  theme_cwd()
```

## Recipe 1: prompt and pitfall

### Adaptation prompt

> Add an explicit `focus` variable that is `"Focus"` for India and `"Context"` otherwise. Map it to `fill` and use a named manual scale. Hide the legend. Do not rely on row or factor order.

### Common generated mistake

```r
scale_fill_manual(values = c("grey80", "#006DAE"))
```

Without names, colours are matched using the scale's ordering. The code can silently emphasise the wrong category after a change in the data.

## Recipe 2: draw context first, then emphasis

### Problem

A manual scale changes colour, but the focal line can still be hidden underneath other lines.

```r
ggplot(housing, aes(date, median, group = city)) +
  geom_line(colour = "grey75", linewidth = 0.7) +
  geom_line(
    data = filter(housing, city == "Austin"),
    colour = "#006DAE", linewidth = 1.4
  )
```

. . .

Layer order is drawing order: context first, focus second.

## Austin sits within a wider Texas housing market

```{r}
#| label: highlight-one-line
#| echo: false
#| fig-height: 4.2
ggplot(housing, aes(date, median, group = city)) +
  geom_line(colour = "grey75", linewidth = 0.75) +
  geom_line(
    data = filter(housing, city == "Austin"),
    colour = brand[["accent"]], linewidth = 1.5
  ) +
  scale_y_continuous(labels = label_dollar(scale = 1e-3, suffix = "k")) +
  labs(
    x = NULL, y = "Median sale price",
    title = "Austin sits within a wider Texas housing market",
    subtitle = "The focal series is drawn last so it remains visible",
    caption = "Source: ggplot2::txhousing"
  ) +
  theme_cwd()
```

## Recipe 2: prompt and pitfall

### Adaptation prompt

> Keep every city visible in light grey, then draw Austin again in blue with a thicker line. Use two `geom_line()` layers so Austin is always on top. Do not remove the comparison cities.

### Common generated mistake

```r
filter(housing, city == "Austin") |>
  ggplot(aes(date, median)) +
  geom_line()
```

Filtering the plot data removes the context. Emphasis is a relationship between foreground and background.

## Recipe 3: replace legend lookup with end labels

```r
endpoints <- housing |>
  group_by(city) |>
  slice_max(date, n = 1, with_ties = FALSE) |>
  ungroup()

ggrepel::geom_text_repel(
  data = endpoints,
  aes(label = city, colour = focus),
  direction = "y", hjust = 0, nudge_x = 90,
  segment.colour = NA, seed = 5523
)
```

. . .

Build a label dataset explicitly. Do not ask a text layer to label every observation.

## End labels keep names beside the evidence

```{r}
#| label: end-labels
#| echo: false
#| fig-height: 4.3
line_colours <- c("Context" = "grey48", "Focus" = brand[["accent"]])

ggplot(housing, aes(date, median, group = city)) +
  geom_line(aes(colour = focus, linewidth = focus)) +
  geom_text_repel(
    data = housing_end,
    aes(label = city, colour = focus),
    direction = "y", hjust = 0, nudge_x = 90,
    segment.colour = NA, seed = 5523, show.legend = FALSE
  ) +
  scale_colour_manual(values = line_colours, guide = "none") +
  scale_linewidth_manual(
    values = c("Context" = 0.7, "Focus" = 1.4), guide = "none"
  ) +
  scale_x_date(expand = expansion(mult = c(0.01, 0.20))) +
  scale_y_continuous(labels = label_dollar(scale = 1e-3, suffix = "k")) +
  coord_cartesian(clip = "off") +
  labs(
    x = NULL, y = "Median sale price",
    title = "End labels keep names beside the evidence",
    caption = "Source: ggplot2::txhousing"
  ) +
  theme_cwd() +
  theme(legend.position = "none")
```

## Recipe 3: prompt and pitfall

### Adaptation prompt

> Replace the legend with one label at the final date for each city. Construct an endpoint data frame, allow labels to move only vertically, add space on the right, and turn clipping off. Keep Austin blue and other labels grey.

### Common generated mistake

```r
geom_text(aes(label = city))
```

This draws a label at every observation. Filter the label layer to one endpoint per group.

## Recipe 4: annotate evidence, not decoration

```r
reference <- mean(airport_year$mean_max)
highest <- slice_max(airport_year, mean_max, n = 1)

ggplot(airport_year, aes(year, mean_max)) +
  geom_hline(yintercept = reference) +
  geom_line() +
  geom_point(data = highest) +
  geom_label_repel(
    data = highest,
    aes(label = sprintf("%d: %.1f°C", year, mean_max))
  )
```

. . .

Use a reference line to establish context and a selective annotation to explain the exceptional observation.

## One annotation directs attention to the exceptional year

```{r}
#| label: selective-annotation
#| echo: false
#| fig-height: 4.2
ggplot(airport_year, aes(year, mean_max)) +
  geom_hline(
    yintercept = airport_reference,
    colour = "grey65", linetype = "dashed"
  ) +
  geom_line(colour = "grey45", linewidth = 0.8) +
  geom_point(data = airport_high, colour = brand[["accent"]], size = 3) +
  geom_label_repel(
    data = airport_high,
    aes(label = sprintf("%d: %.1f°C", year, mean_max)),
    colour = brand[["accent"]], fill = "white",
    seed = 5523, min.segment.length = 0
  ) +
  scale_x_continuous(breaks = breaks_width(10)) +
  scale_y_continuous(labels = label_number(suffix = "°C")) +
  labs(
    x = NULL, y = "Temperature (°C)",
    title = "One annotation directs attention to the exceptional year",
    subtitle = "Dashed line: series mean",
    caption = "Source: Bureau of Meteorology — Melbourne Airport"
  ) +
  theme_cwd() +
  theme(plot.caption.position = "panel")
```

## Recipe 4: prompt and pitfall

### Adaptation prompt

> Add a dashed reference line at the series mean. Highlight and label only the year with the highest annual mean. Format the y-axis in degrees Celsius and state what the dashed line means in the subtitle.

### Common generated mistake

```r
geom_text(aes(label = round(mean_max, 1)))
```

Labelling every value converts the plot into a crowded table. Annotation should answer a particular question.

## Recipe 5: make the finishing layer reusable

```r
theme_cwd <- function(base_size = 16) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title.position = "plot",
      plot.caption.position = "plot",
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.title = element_text(face = "bold"),
      plot.caption = element_text(hjust = 0),
      legend.position = "bottom",
      plot.margin = margin(10, 36, 10, 10)
    )
}
```

One function makes deliberate defaults easy to reuse and review.

## Recipe 5: prompt and pitfall

### Adaptation prompt

> Extract the non-data styling from this plot into a function called `theme_cwd(base_size = 16)`. Start from `theme_minimal()`. Do not set data colours or axis scales inside the theme.

### Common generated mistake

```r
theme_set(theme_cwd())
```

Changing the session-wide theme can alter unrelated plots. Add `theme_cwd()` to each plot unless a global change is intentional.

## A polished plot still needs a review

::: {.columns}
::: {.column width="48%"}

### Evidence

* Does the title match the data shown?
* Are the denominator and time period clear?
* Does emphasis hide relevant context?
* Are annotations computed from the data?

:::
::: {.column width="48%"}

### Communication

* Is the intended comparison easy?
* Is colour doing a specific job?
* Are labels legible at output size?
* Can the plot work without its legend?

:::
:::

## Week 7 lesson

::: {.callout-important}

## Summary

* CSS selectors scope a visual change; declarations describe the change.
* A good LLM brief names the target, context, constraints, and acceptance check.
* Generated code is a draft to render, inspect, and revise.
* Explicit focus variables, deliberate layer order, direct labels, and selective annotations create hierarchy in `ggplot2`.
* Reusable palettes and themes make consistency easier without replacing judgement.

:::

## Resources {.smaller}

* [Quarto: HTML theming and CSS](https://quarto.org/docs/output-formats/html-themes.html)
* [MDN: CSS first steps](https://developer.mozilla.org/en-US/docs/Learn_web_development/Core/Styling_basics)
* [`ggplot2`: manual scales](https://ggplot2.tidyverse.org/reference/scale_manual.html)
* [`ggplot2`: theme elements](https://ggplot2.tidyverse.org/reference/theme.html)
* [`ggrepel`: repulsive text annotations](https://ggrepel.slowkow.com/reference/geom_text_repel.html)

::: notes
[Sources]

- https://quarto.org/docs/output-formats/html-themes.html
- https://developer.mozilla.org/en-US/docs/Learn_web_development/Core/Styling_basics
- https://ggplot2.tidyverse.org/reference/scale_manual.html
- https://ggplot2.tidyverse.org/reference/theme.html
- https://ggrepel.slowkow.com/reference/geom_text_repel.html
:::
