---
title: "ETC5523: Communicating with Data"
---

```{r setup, include = FALSE}
# please add any R-packages used here in DESCRIPTION imports
library(yaml)
library(tibble)
library(dplyr)
library(purrr)
library(glue)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE)
```

```{r fns}
make_rectangular <- function(.data) {
  map_dfr(.data, ~ tibble(!!!.x))
}


ui_list <- function(vec) {
  for (x in vec) {
    cat("-", x, "\n")
  }
}
```


```{r staffs}
dat <- read_yaml("_course.yml")
team_df <- make_rectangular(dat$team)
schedule_df <- make_rectangular(dat$schedule)

nlecturer <- sum(team_df$role == "lecturer")
ntutor <- sum(team_df$role == "tutor")
```


## Lecturer`r ifelse(nlecturer > 1, 's', '')`

```{r lecturers, results="asis"}
team_df %>%
  filter(role == "lecturer") %>%
  pull(name) %>%
  ui_list()
```

## Tutor`r ifelse(ntutor > 1, 's', '')`

```{r tutors, results="asis"}
team_df %>%
  filter(role == "tutor") %>%
  pull(name) %>%
  ui_list()
```

## Consultations

All consultations will be in using zoom. Check Moodle for the links.

```{r consult, results="asis"}
team_df %>%
  mutate(name = gsub(" (Chief Examiner)", "", name, fixed = T)) %>%
  mutate(consult_print = glue::glue("**{name}**: {consult}")) %>%
  pull(consult_print) %>%
  ui_list()
```


## Schedule

* Lectures: **Wed 6.00-8.00pm** F2F - See your timetable for the room
* Tutorial A: **Thu 7.30-9.00pm** F2F - See your timetable for the room
* Tutorial B: **Fri 6.00-7.30pm** F2F - See your timetable for the room
* Tutorial C: **Fri 7.30-9.00pm** F2F - See your timetable for the room

There are no lectures or tutorials during the midsemester break.

```{r}
schedule_df <- schedule_df %>%
  mutate(
    lecture = file.exists(sprintf("lectures/lecture-%.2d.html", week)),
    lecture_a = file.exists(sprintf("lectures/lecture-%.2dA.html", week)),
    lecture_b = file.exists(sprintf("lectures/lecture-%.2dB.html", week)),
    tutorial = file.exists(sprintf("tutorials/tutorial-%.2d-question.html", week)),
    solution = file.exists(sprintf("tutorials/tutorial-%.2d-solution.html", week))
  ) %>%
  mutate(
    slides = case_when(
      lecture_a & lecture_b ~ glue("A: <a class='fa fa-link' href='lectures/{sprintf(\"lecture-%.2dA.html\", week)}' target='_blank'></a> <a class='fa fa-file-pdf' href='lectures/{sprintf(\"lecture-%.2dA.pdf\", week)}' target='_blank'></a><br>B: <a class='fa fa-link' href='lectures/{sprintf(\"lecture-%.2dB.html\", week)}' target='_blank'></a> <a class='fa fa-file-pdf' href='lectures/{sprintf(\"lecture-%.2dB.pdf\", week)}' target='_blank'></a>"),
      lecture ~ glue("<a class='fa fa-link' href='lectures/{sprintf(\"lecture-%.2d.html\", week)}' target='_blank'></a> <a class='fa fa-file-pdf' href='lectures/{sprintf(\"lecture-%.2d.pdf\", week)}' target='_blank'></a>"),
      TRUE ~ ""
    ),
    tut = case_when(
      tutorial & solution ~ glue("<a class='fa fa-tasks' href='tutorials/{sprintf(\"tutorial-%.2d\", week)}-question.html' target='_blank'></a>  <a class='fa fa-lightbulb' href='tutorials/{sprintf(\"tutorial-%.2d\", week)}-solution.html' target='_blank'></a>"),
      tutorial ~ glue("<a class='fa fa-tasks' href='tutorials/{sprintf(\"tutorial-%.2d\", week)}-question.html' target='_blank'></a>"),
      TRUE ~ ""
    ),
    topic = ifelse(is.na(topic), "", topic),
    readings = ifelse(is.na(readings), "", readings),
    assessment = ifelse(is.na(assessment), "", assessment),
    tools = ifelse(is.na(tools), "", tools)
  ) %>%
  select(week, slides, tut, topic, readings, assessment)


midsem_break <- 9 + 1 # after which week

schedule_df %>%
  kable(
    escape = FALSE, align = c("c", "c", "c", "c", "l", "l", "l", "l"),
    col.names = c(
      "Week", "Slides", "Tutorial",
      "Topic", "Resources", "Assessments"
    )
  ) %>%
  kable_styling(
    bootstrap_options = "striped",
    fixed_thead = T
  ) %>%
  column_spec(1, width = "0.5cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "1cm") %>%
  group_rows("Midsemester Break (1 week)", midsem_break + 1, 12, label_row_css = "background-color: #d9e1e8;") %>%
  add_indent(1:midsem_break) # this needs to be done to match up with above
```

