---
title: "ETC5523: Communicating With Data"
toc: false
format:
  html:
    page-layout: full
---

This unit develops teamwork skills, through instruction on reproducibility and version control to conduct collaborative data analysis. These skills are practised in a team setting with a data analysis project.

## Learning outcomes

1. Effectively communicate data analysis, using a blog, reports and presentations
2. Learn how to build a web app to provide an interactive data analysis
3. Construct a data story

## Teachers

* [**Michael Lydeamore**](https://www.michaellydeamore.com). Chief Examiner. Email: [michael.lydeamore@monash.edu](mailto:michael.lydeamore@monash.edu)
* [**Maliny Po**]. Email: [maliny.po@monash.edu](mailto:maliny.po@monash.edu)

```{r}
#| label: load_packages
#| include: false
#| message: false
#| warning: false
#| echo: false
#| cache: false
library(tidyverse)
options(knitr.kable.NA = '')
source(here::here("course_info.R"))
week <- as.integer(1 + ((Sys.Date() - as_date(start_semester))/7))
```

## Weekly schedule

* 1 hour workshop, Mondays 10am-11am each week
* 2 hour seminar, Monday 6pm-8pm each week
* 1.5 hour tutorial, one of:
  * Monday 12pm-1pm
  * Monday 1pm-2pm
  * Monday 2pm-3pm

## Consultation

* Michael Lydeamore: Tuesdays 1.30pm-2.30pm (Rm 354, Building 6)
* Maliny Po: Wednesday 1pm-2.30pm (Rm 232A, Building 6)

```{r}
#| label: schedule2
#| message: false
#| warning: false
#| echo: false
#| output: asis
schedule |>
    transmute(
        Week = if_else(Topic == "Mid-semester break", NA, Week, missing = Week),
        Date = format(Date, "%d %b"),
        Show = (!is.na(Week) & Week <= 12) & !grepl("University Holiday", Topic) & file.exists(glue::glue("./week{Week}/index.qmd")),
        Show_Tutorial = !is.na(Week) & Week <= 12,
        Topic = if_else(!Show, Topic, glue::glue('[{{Topic}}](./week{{Week}}/index.html){target="_blank"}', .open = "{{", .close = "}}")),
        Tutorial = if_else(!Show_Tutorial, "", glue::glue('[{{< fa list-check >}}](./week{{{Week}}}/tutorial/index.html){target="_blank"}', .open = "{{{", .close = "}}}")),
        Reference = if_else(!Show, Reference, glue::glue("[{Reference}]({Reference_URL})")),
        Assessments = if_else(is.na(Assignment), Assignment, glue::glue("[{Assignment}]({File})"))
    ) |>
    select(-Show, -Show_Tutorial) |>
    knitr::kable(format = "markdown")
```