---
title: 'ETC5523: Communicating with Data'
subtitle: "Tutorial 8"
author: "Michael Lydeamore"
date: "Week 8"
format: 
  unilur-html:
    output-file: index.html
  unilur-html+solution:
    output-file: solution.html
---


```{r setup, include = FALSE, eval = file.exists("setup.R")}
library(tidyverse)
```

## 🎯 Objectives


* make your first R package
* make a simple R package as a container for data

::: {.callout-note collapse="true"}

## Preparation

```{r}
#| eval: false

install.packages(c("devtools","usethis","roxygen2"))
```

:::

    
##  Exercise 8A

Use the `usethis` package to create a new package called `cwdata`

::: {.unilur-solution}

```{r}
#| eval: false
library(usethis)

create_package("cwdata")
```

:::


**Edit the DESCRIPTION file** created and modify the `Authors@R` tag to include yourself as the creator and author of the package. 


##  Exercise 8B

Use the `usethis` package to create an R script called `key_crop_yields.R` in the  `data-raw` directory in your packages folder. 

::: {.unilur-solution}

```{r}
#| eval: false
use_data_raw(name = "key_crop_yields")
```

:::


The `data-raw` directory is used to hold scripts that generate package data. We will use the data from [Our World in Data](https://ourworldindata.org/crop-yields) and sourced from [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-01/readme.md). This data contains agricultural yields across crop types and by entity (country or region) from 1960 to 2018. 

**Edit the script** to process the data into a long form tidy representation. Here's some code to get you started:

```{r data-raw}
#| echo: true
#| eval: false

library(janitor)
library(tidyverse)
url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-01/key_crop_yields.csv'

key_crop_yields <- read_csv(url) %>% 
  clean_names() %>% 
  pivot_longer(
    cols = -___,
    names_to = "crop", 
    values_to = "yield",
    names_pattern = "([^_]+)"
  )
```

After you've finished writing your script make sure you've run it.

You should now have a folder called `data` in your package,
with a file called `key_crop_yields.rda`.


## Exercise 8C

We now need to create an R script that lives in the `R` directory
that documents our data object `key_crop_yields`.

```{r}
#| eval: false
#| echo: true
use_r("key_crop_yields.R")
```

Modify the script and **document the data.**

- Describe the number of rows and columns 
- Type of object `key_crop_yields` is. 
- What are the types of the columns?
- What do the columns mean?
- Where did you find it? 

It is possible to use ordinary markdown syntax in documenation with `roxygen2` by running (in fact this is set by default if you used `usethis::create_package()`) 

```{r}
#| eval: false
#| echo: true
use_roxygen_md()
```

Once you've finished modifying run

```{r}
#| echo: true
#| eval: false
devtools::document()
```

at the console to build your package documentation.

## Exercise 8D

We now have a minimally working package that contains some data. 
Restart your R session and then run 

```{r}
#| eval: false
#| echo: true
devtools::load_all()
```

Now the `key_crop_yields` data is your available to your R session.

**Edit your DESCRIPTION file** to have a more appropriate title and description and version number. 

After you've finished editing run:
```{r}
#| eval: false
#| echo: true
devtools::check()
```

**Do you get any errors or warnings?**

Once you have passed the check, install the package locally:

```{r}
#| eval: false
#| echo: true
devtools::install()
```

Restart your R session and try loading your package. Check out the help page for the data.

::: {.unilur-solution}

```{r}
#| eval: false
library(cwdata)
key_crop_yields
?key_crop_yields
```

:::


