---
title: 'ETC5523: Communicating with Data'
subtitle: "Tutorial 8"
author: "Michael Lydeamore"
format: 
  unilur-html:
    output-file: index.html
  unilur-html+solution:
    output-file: solution.html
---


```{r setup, include = FALSE, eval = file.exists("setup.R")}
library(tidyverse)
```

## 🎯 Objectives

By the end of this tutorial, you will be able to:

* create your first R package;
* prepare and include a dataset reproducibly;
* document package data with `roxygen2`; and
* check, install and use the finished package.

::: {.callout-note collapse="true"}

## Preparation

```{r}
#| eval: false

install.packages(c("devtools", "janitor", "roxygen2", "tidyverse", "usethis"))
```

:::

    
##  Exercise 8A

Use `usethis` to create a new package called `trashwheeldata`.

::: {.unilur-solution}

```{r}
#| eval: false
usethis::create_package("trashwheeldata")
```

:::


**Edit `DESCRIPTION`** and replace the placeholders with:

* yourself as the creator and author;
* a meaningful title and description; and
* an appropriate version number.

Add an MIT licence, replacing the placeholder name with your own:

```{r}
#| eval: false
#| echo: true

usethis::use_mit_license("Your Name")
```


##  Exercise 8B

Use `usethis` to create an R script called `trash_collections.R` in the `data-raw` directory of your package.

::: {.unilur-solution}

```{r}
#| eval: false
usethis::use_data_raw("trash_collections")
```

:::


The `data-raw` directory holds scripts that generate package data. We will use the [Trash Wheel Collection Data](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-03-05/readme.md) published through TidyTuesday and sourced from Baltimore's [Healthy Harbor initiative](https://www.mrtrashwheel.com/).

Four semi-autonomous Trash Wheels collect rubbish flowing into Baltimore's waterways. The data record each dumpster load, including its weight and volume and estimated counts of several kinds of rubbish.

**Edit the script** to create a long-form dataset in which each row represents one rubbish category from one dumpster collection. Complete the missing `cols` expression.

```{r data-raw}
#| echo: true
#| eval: false

url <- paste0(
  "https://raw.githubusercontent.com/rfordatascience/",
  "tidytuesday/main/data/2024/2024-03-05/trashwheel.csv"
)

trash_collections <- readr::read_csv(url, show_col_types = FALSE) |>
  janitor::clean_names() |>
  dplyr::mutate(date = lubridate::mdy(date)) |>
  tidyr::pivot_longer(
    cols = ___,
    names_to = "trash_type",
    values_to = "count"
  )

usethis::use_data(trash_collections, overwrite = TRUE)
```

::: {.unilur-solution}

The seven item-count columns run from `plastic_bottles` to `sports_balls`:

```{r}
#| eval: false
#| echo: true

trash_collections <- readr::read_csv(url, show_col_types = FALSE) |>
  janitor::clean_names() |>
  dplyr::mutate(date = lubridate::mdy(date)) |>
  tidyr::pivot_longer(
    cols = plastic_bottles:sports_balls,
    names_to = "trash_type",
    values_to = "count"
  )

usethis::use_data(trash_collections, overwrite = TRUE)
```

:::

Run the completed script from beginning to end.

You should now have a folder called `data` in your package,
with a file called `trash_collections.rda`.

::: {.callout-tip}
The script in `data-raw/` is the reproducible source. The `.rda` file in `data/` is the object shipped to users.
:::

How many rows do you expect after pivoting seven rubbish categories? Confirm your prediction with `dim(trash_collections)`.

::: {.unilur-solution}

The source has 993 dumpster collections. Pivoting seven categories produces $993 \times 7 = 6{,}951$ rows and 11 columns.

:::


## Exercise 8C

Create an R script in the `R` directory to document `trash_collections`.

```{r}
#| eval: false
#| echo: true
usethis::use_r("trash_collections")
```

Modify the script and **document the data.**

- Give the dataset a concise title and description.
- Describe its dimensions and object type.
- Define every column, including its units.
- Explain that item counts are estimates and that missing counts are not necessarily zero.
- Cite both TidyTuesday and the original Healthy Harbor source.
- Finish the block with the quoted object name, `"trash_collections"`.

Markdown support is enabled by default for packages created with `usethis::create_package()`. You can confirm or enable it with:

```{r}
#| eval: false
#| echo: true
usethis::use_roxygen_md()
```

Once you've finished modifying run

```{r}
#| echo: true
#| eval: false
devtools::document()
```

at the console to build your package documentation.

::: {.unilur-solution}

One possible documentation block is:

```r
#' Rubbish collected by Baltimore's Trash Wheel family
#'
#' Dumpster collections from four Trash Wheels operating in Baltimore's
#' waterways. Each row describes one rubbish category in one collection.
#' Item counts are estimated using samples taken during collection.
#'
#' @format A data frame with 6,951 rows and 11 variables:
#' \describe{
#'   \item{id}{Short identifier for the Trash Wheel.}
#'   \item{name}{Name of the Trash Wheel.}
#'   \item{dumpster}{Dumpster collection number.}
#'   \item{month}{Month of collection.}
#'   \item{year}{Year of collection.}
#'   \item{date}{Date of collection.}
#'   \item{weight}{Weight of collected rubbish, in US tons.}
#'   \item{volume}{Volume of collected rubbish, in cubic yards.}
#'   \item{homes_powered}{Source estimate of homes that could be powered
#'     using energy generated from the collected rubbish.}
#'   \item{trash_type}{Category of rubbish counted.}
#'   \item{count}{Estimated item count; missing when not reported.}
#' }
#' @source TidyTuesday,
#'   \url{https://github.com/rfordatascience/tidytuesday/tree/main/data/2024/2024-03-05};
#'   original data from Baltimore Healthy Harbor,
#'   \url{https://www.mrtrashwheel.com/}.
"trash_collections"
```

:::

## Exercise 8D

We now have a minimally working package containing a documented dataset. Restart your R session and run:

```{r}
#| eval: false
#| echo: true
devtools::load_all()
```

Confirm that `trash_collections` is available and that its help page answers the questions below:

```{r}
#| eval: false
#| echo: true

trash_collections
?trash_collections
```

* What does one row represent?
* Are the item counts observed exactly or estimated?
* Does a missing count mean that no items were collected?
* What units are used for `weight` and `volume`?

Then check the complete package:
```{r}
#| eval: false
#| echo: true
devtools::check()
```

Fix every error and warning. Read each note and decide whether it requires action.

Once you have passed the check, install the package locally:

```{r}
#| eval: false
#| echo: true
devtools::install()
```

Restart your R session, load the installed package, and open the help page again.

::: {.unilur-solution}

```{r}
#| eval: false
library(trashwheeldata)
trash_collections
?trash_collections
```

:::

## Completion check

You have finished when:

* `data-raw/trash_collections.R` reproduces the packaged data;
* `data/trash_collections.rda` exists;
* `devtools::document()` creates the help page;
* `DESCRIPTION` and the licence contain no placeholders;
* `devtools::check()` reports no errors or warnings; and
* the installed package works in a clean R session.

