---
title: 'ETC5523: Communicating with Data'
subtitle: "Tutorial 3"
author: "Michael Lydeamore"
date: "Week 3"
format:
    unilur-html:
        output-file: index.html
        embed-resources: true
    unilur-html+solution:
        output-file: solution.html
        embed-resources: true
---


```{r setup, include = FALSE, eval = file.exists("setup.R")}
source("setup.R")
```

```{r}
#| echo: false
#| eval: !expr file.exists("tutorial.css")
htmltools::includeCSS("tutorial.css")
```

```{r}
#| include: false
checkbox <- function(text) {
  paste0('<label class="checkbox-container"><input type="checkbox"> ', text, '<span class="checkmark"></span>
</label>')
}
```

## 🎯 Objectives

* use Quarto for creating reproducible documents
* make blog websites using Quarto
* share webpage and websites on public domains with Quarto Pubs, GitHub Pages or Netlify

::: {.callout-note collapse="true"}

# Preparation

1. Install the [latest RStudio IDE](https://www.rstudio.com/products/rstudio/download/)
2. Install the [latest Quarto CLI](https://quarto.org/docs/get-started/)
3. (Optional) Make a (free) account at [https://quartopub.com/](https://quartopub.com/) if you want to share your website via Quarto Pub.
4. (Optional) Make a (free) account at [https://www.netlify.com/](https://www.netlify.com/) if you want to share your website via Netlify. 

:::

Note: there will be no tutorial solution this week as the task is about getting you set up for making a website using Quarto. 


## 🌐️ Exercise 4A

**Construct a data story using Quarto**

Convert your assignment 1 into a Quarto document. The YAML below may be helpful:

``` yaml
---
title: "My story title"
author: "Your name"
format: html
---
```

`r checkbox("Write this story in a qmd file with a HTML output")`    
`r checkbox("Render this qmd file")`  


## 🌐️ Exercise 4B

**Make a blog website using Quarto**

Create a blog website with Quarto, as shown in lecture recording. Add the data story you constructed in Exercise 4A as a blog post. 

`r checkbox("Make a Quarto blog website")`  
`r checkbox("Add one blog post")`  


## 🌐️ Exercise 4C

**Publish your blog website**

Publish your blog website constructed in Exercise 4B using either Quarto Pub, GitHub Pages or Netlify.

`r checkbox("Upload this to a GitHub repo")`  
`r checkbox("Publish this on the web")`
