---
title: 'ETC5523: Communicating with Data'
subtitle: "Tutorial 2"
author: "Michael Lydeamore"
date: "Week 2"
format:
    unilur-html:
        output-file: index.html
        embed-resources: true
    unilur-html+solution:
        output-file: solution.html
        embed-resources: true
---


```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```


## `r emo::ji("target")` Objectives

- Understand structure of journalism articles
- Add data to a story to improve the impact

## Introduction 

You will be analysing an article from BBC World News about Japan's ageing population. The article is available online on the BBC website [here](https://www.bbc.com/news/world-asia-66850943).

Form small groups (3-5 people) for these exercises.

## 👥 Exercise 1

a. In the article, highlight each section of the "inverted pyramid" model. Note that not all parts of the inverted pyramid may be present in the article.

::: {.unilur-solution}
**Main facts:** First two paragraphs include the main facts

**More detail:** Followed by details of _why_ this may be the case.

**Quote::** From Japanese Prime MInister Fumio Kishida.

There's no bio, and little additional detail, likely because this is a very short article and so there was no space to include the extra information.
:::


b. Is this article a news feature or a news story? Why?

::: {.unilur-solution}
This is a news story: It is quoting sources and repeating information. There is no inclusion of characters in the story.
:::

c. Try and write a new headline for the article that conveys a similar message. Remember the principles of concise writing when developing your new headline.

::: {.unilur-solution}
There are lots of potential approaches to this. Here is one:

Japan's ageing population: Now the oldest in the world
:::


## Exercise 2

With a good understanding of the content of the article, in your groups, you are tasked with finding some data to enrich this story, and producing a graphic to add to the article.

You will need to:

a. Locate a suitable data source,

b. Write code to load in the data and manipulate it into a format appropriate for visualisation,

c. Create a clear, concise graphic that supports the message of the story.

::: {.unilur-solution}
You could pick a few aspects of the story to look at data from. I've chosen the population data, which I've sourced from [Our World in Data](https://ourworldindata.org/population-growth).

```{r}
#| echo: true
#| fig-height: 6

library(tidyverse)
library(viridisLite)
library(ggtext)

population_data <- read_csv("population-and-demography.csv")

plot_data <- population_data |>
    # The article mentions Italy, Finland and Japan
    filter(Code %in% c("ITA", "FIN", "JPN")) |>
    select(!Code) |>
    pivot_longer(!c(Entity, Year)) |>
    group_by(Entity, Year) |>
    # The populations of these countries are very different, so we should normalised them to put them on the same scale
    mutate(normalised_value = value / sum(value)) |>
    ungroup() |>
    # We don't want them in alphabetical order: Japan is the takeaway country
    mutate(Entity = factor(Entity, levels = c("Japan", "Italy", "Finland"))) |>
    # Only old populations
    filter(name == "65+ years") |>
    filter(Year >= 1980)

# One trick is to use text labels instead of a legend to draw the reader to a key part of the plot.
# This puts the labels on the right hand side (max(year)) and at the correct height.
labels <- plot_data |>
    group_by(Entity) |>
    summarise(max_year = max(Year), y_pos = normalised_value[which.max(Year)])

colors <- c("#2600ff", "#a38687", "#a38687")

ggplot(plot_data, aes(x=Year, y = normalised_value*100, colour = Entity)) +
    geom_line() +
    geom_point() +
    geom_text(aes(label = Entity, x = max_year+1, y = y_pos*100), data = labels, hjust = 0) +
    labs(x = "Year", y = "Percentage of population", title = "<span style ='color:#2600ff;'>Japan</span> has the world's oldest population") + 
    guides(colour = "none") +
    # Space for labels
    expand_limits(x = 2030) +
    # We are only colouring one line - the one we want people to look at!
    scale_color_manual(values = colors) +
    cowplot::theme_cowplot() +
    theme(plot.title = element_markdown()) +
    annotate(geom = "text", x = 2030, y = -4, label = "Data source: Our World in Data", size = 2, hjust = 1) +
    coord_cartesian(ylim = c(0, 30), clip = "off")
```
:::


## Exercise 3

Present your visualisation to the class. Explain why you chose a particular data source, and what the audience can learn from the visualisation.

::: {.callout-tip}
Try to ensure your graphic _adds_ to the story. You want the visualisation to give extra information to the audience.
:::

::: {.unilur-solution}
This visualisation shows people how the age of Japan's population has _changed_ over time, which is not otherwise present in the article. It also shows the scale of the difference between Japan and two other European countries.

It could be improved: There are too many years on the x-axis, the y-axis labels should probably have % signs on them, and the title could be better.

For the data source annotation, best practice would be to include a link, although this isn't generally possible inside `ggplot`, so we instead include a plain text reference to the website.
:::


