---
pagetitle: "ETC5523: Communicating with Data"
subtitle: "One dataset, three stories"
author: "Michael Lydeamore"
email: "michael.lydeamore@monash.edu"
date: "Week 10"
department: "Department of Econometrics and Business Statistics"
unit-url: "cwd.numbat.space"
footer: "ETC5523 Week 10"
format:
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: c
    multiplex: false
    theme: ../assets/monash.scss
    show-slide-number: all
    show-notes: true
    controls: true
    width: 1280
    height: 720
    css: [../assets/tachyons-addon.css, ../assets/custom.css]
    include-after-body: "../assets/after-body.html"
    incremental: true
execute:
  echo: true
  eval: true
---

```{r}
#| include: false
library(dplyr)
library(ggplot2)
library(lubridate)
library(readr)
library(tidyr)

solar_demo <- read_csv("tutorial/solar_data.csv", show_col_types = FALSE) |>
  filter(con_gen == "Consumption") |>
  mutate(
    datetime = as.POSIXct(datetime, tz = "UTC"),
    date = as.Date(datetime),
    hour = hour(datetime) + minute(datetime) / 60,
    daylight = pmax(sin(pi * (hour - 6) / 12), 0),
    seasonal = 0.70 + 0.30 * cos(2 * pi * (yday(datetime) - 15) / 365.25),
    solar_kwh = 5 * daylight^1.5 * seasonal * 0.5
  )

dispatch_demo <- function(data, capacity_kwh = 10, round_trip_efficiency = 0.90) {
  one_way_efficiency <- sqrt(round_trip_efficiency)
  battery_state <- grid_import <- grid_export <- numeric(nrow(data))
  state <- 0

  for (i in seq_len(nrow(data))) {
    net <- data$net_demand_kwh[i]

    if (net < 0) {
      surplus <- -net
      stored <- min(surplus * one_way_efficiency, capacity_kwh - state)
      state <- state + stored
      grid_export[i] <- surplus - stored / one_way_efficiency
    } else {
      delivered <- min(net, state * one_way_efficiency)
      state <- state - delivered / one_way_efficiency
      grid_import[i] <- net - delivered
    }

    battery_state[i] <- state
  }

  data |>
    mutate(
      battery_state_kwh = battery_state,
      grid_import_kwh = grid_import,
      grid_export_kwh = grid_export
    )
}

trace_day <- solar_demo |>
  filter(date == as.Date("2024-12-06")) |>
  mutate(net_demand_kwh = energy_kwh - solar_kwh) |>
  dispatch_demo()
```

## <br>[`r rmarkdown::metadata$pagetitle`]{.monash-blue} {#etc5523-title background-image="images/bg-01.png"}

### `r rmarkdown::metadata$subtitle`

Lecturer: *`r rmarkdown::metadata$author`*

`r rmarkdown::metadata$department`

::: tl
<br>

<ul class="fa-ul">
<li>[<i class="fas fa-envelope"></i>]{.fa-li}`r rmarkdown::metadata$email`</li>
<li>[<i class="fas fa-calendar-alt"></i>]{.fa-li} `r rmarkdown::metadata$date`</li>
<li>[<i class="fa-solid fa-globe"></i>]{.fa-li}<a href="`r rmarkdown::metadata[["unit-url"]]`">`r rmarkdown::metadata[["unit-url"]]`</a></li>
</ul>
:::

## By the end of today

You will be able to:

1. build one shared reactive model that feeds several views;
2. use tabs, conditional UI and reactive text to guide an argument; and
3. distinguish legitimate emphasis from changing or hiding the evidence.

::: {.callout-important .fragment}
The technical goal is not three separate apps. It is **one calculation, communicated three ways**.
:::

## The same evidence can support different stories

::: {.columns}
::: {.column width="33%"}
### Salesperson

“A battery helps you use more of the energy your panels generate.”
:::

::: {.column width="33%"}
### Neutral analyst

“Savings depend on production, demand, tariffs and timing.”
:::

::: {.column width="33%"}
### Sceptic

“Up-front costs remain visible long after the first bill saving.”
:::
:::

::: {.fragment}
All three statements can be true at the same time.
:::

## First establish the common evidence

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 4.3
#| out-width: "88%"
solar_demo |>
  filter(date >= as.Date("2024-12-01"), date <= as.Date("2024-12-07")) |>
  select(datetime, Consumption = energy_kwh, `Modelled solar` = solar_kwh) |>
  pivot_longer(-datetime, names_to = "series", values_to = "energy_kwh") |>
  ggplot(aes(datetime, energy_kwh, colour = series)) +
  geom_line(linewidth = 0.7) +
  scale_colour_manual(values = c("#006DAE", "#F28E2B")) +
  labs(x = NULL, y = "Energy per half-hour (kWh)", colour = NULL) +
  theme_minimal(base_size = 18) +
  theme(legend.position = "top")
```

Consumption is observed. Solar production is a transparent, simplified model.

# Build the scenario one column at a time

## Begin with energy measured every half-hour

The supplied data are already in long form:

~~~r
consumption_data
# datetime              energy_kwh
# 2024-07-15 00:00:00       0.0312
# 2024-07-15 00:30:00       0.0437
~~~

~~~r
interval_hours <- median(
  as.numeric(diff(consumption_data$datetime), units = "hours")
)
# 0.5
~~~

The interval length is needed because **kW measures power**, while each row requires **kWh of energy**.

## The solar model converts capacity into interval energy {.smaller}

For each half-hour:

$$
\text{solar kWh} =
\text{system kW} \times
\text{daylight shape} \times
\text{seasonal factor} \times
\text{interval hours}
$$

~~~r
hour <- hour(datetime) + minute(datetime) / 60
daylight <- pmax(sin(pi * (hour - 6) / 12), 0)
seasonal <- 0.70 + 0.30 * cos(2 * pi * (yday(datetime) - 15) / 365.25)
solar_kwh <- system_kw * daylight^1.5 * seasonal * interval_hours
~~~

This is a teaching model, not a site-specific production forecast.

## Inspect each part before combining it

::: {.columns}
::: {.column width="33%"}
### Daylight

Zero outside 6 am–6 pm; highest near midday.
:::

::: {.column width="33%"}
### Season

The factor ranges from 0.4 to 1.0 and peaks in the Australian summer.
:::

::: {.column width="33%"}
### System and interval

A larger system scales production; multiplying by 0.5 converts half-hourly power to energy.
:::
:::

::: {.callout-tip .fragment}
Keep these intermediate columns while developing. Remove them only after the calculation has been checked.
:::

## Net demand determines the direction of flow

~~~r
net_demand_kwh <- energy_kwh - solar_kwh
~~~

::: {.columns}
::: {.column width="50%"}
### Positive net demand

Consumption exceeds solar production.

Energy must come from the battery or grid.
:::

::: {.column width="50%"}
### Negative net demand

Solar production exceeds consumption.

The surplus can charge the battery or be exported.
:::
:::

## Without a battery, split net demand at zero

~~~r
grid_import_without_battery_kwh <- pmax(net_demand_kwh, 0)
grid_export_without_battery_kwh <- pmax(-net_demand_kwh, 0)
~~~

| Net demand | Grid import | Grid export |
|---:|---:|---:|
| 0.8 kWh | 0.8 kWh | 0 kWh |
| −1.2 kWh | 0 kWh | 1.2 kWh |

The sign tells us the direction; **pmax()** keeps both flow columns non-negative.

## A battery makes the calculation sequential {.smaller}

The state of charge at interval *i* depends on interval *i − 1*.

For every row, in time order:

1. calculate net demand;
2. use surplus to charge, subject to capacity and charging losses;
3. use stored energy to meet demand, subject to discharging losses;
4. record the remaining grid import or export; and
5. carry the new state of charge to the next row.

::: {.callout-important .fragment}
This is why battery dispatch cannot be calculated with one independent **mutate()** expression.
:::

## Dispatch one interval at a time {.smaller}

~~~r
one_way_efficiency <- sqrt(round_trip_efficiency)
state <- 0

for (i in seq_len(nrow(data))) {
  net <- data$net_demand_kwh[i]

  if (net < 0) {
    surplus <- -net
    stored <- min(surplus * one_way_efficiency, capacity_kwh - state)
    state <- state + stored
    grid_export[i] <- surplus - stored / one_way_efficiency
  } else {
    delivered <- min(net, state * one_way_efficiency)
    state <- state - delivered / one_way_efficiency
    grid_import[i] <- net - delivered
  }
}
~~~

The square root applies half of the round-trip loss while charging and half while discharging.

## Trace one day through the battery {.smaller}

```{r}
#| echo: false
#| tbl-cap: "Selected intervals from 6 December 2024; 10 kWh battery, 90% round-trip efficiency."
trace_day |>
  filter(format(datetime, "%H:%M") %in% c("00:00", "06:00", "09:00", "12:00", "15:00", "18:00", "21:00")) |>
  transmute(
    Time = format(datetime, "%H:%M"),
    Consumption = round(energy_kwh, 2),
    Solar = round(solar_kwh, 2),
    Net = round(net_demand_kwh, 2),
    Battery = round(battery_state_kwh, 2),
    Import = round(grid_import_kwh, 2),
    Export = round(grid_export_kwh, 2)
  ) |>
  knitr::kable(align = "rrrrrrr")
```

Read across a row, then down the **Battery** column: that is the state carried through time.

## Tariffs turn energy flows into interval costs {.smaller}

~~~r
baseline_cost <- energy_kwh * import_tariff

scenario_cost <-
  grid_import_kwh * import_tariff -
  grid_export_kwh * feed_in_tariff

interval_saving <- baseline_cost - scenario_cost
~~~

::: {.callout-note .fragment}
A feed-in payment reduces the scenario cost. It does not make exported energy “wasted”.
:::

## Summaries come after the interval calculation

~~~r
metrics <- list(
  consumption = sum(energy_kwh),
  production = sum(solar_kwh),
  grid_import = sum(grid_import_kwh),
  grid_export = sum(grid_export_kwh),
  saving = sum(interval_saving),
  self_sufficiency = 1 - sum(grid_import_kwh) / sum(energy_kwh)
)
~~~

Simple payback then compares an annualised saving with up-front cost.

::: {.callout-warning .fragment}
Annualising a short or unusual period can produce an unstable payback estimate.
:::

## Framing changes attention—not the calculation

We may legitimately change:

- the comparison placed in the foreground;
- the time window or summary, if it is clearly labelled;
- annotations, ordering, colour and language; and
- the question the view helps the user answer.

::: {.callout-warning .fragment}
We must not quietly change tariffs, system size, battery assumptions or the underlying records between tabs.
:::

# Build one source of truth

## A shared reactive prevents three versions of reality

```r
scenario_data <- reactive({
  consumption_data |>
    filter(date >= input$date_range[1],
           date <= input$date_range[2]) |>
    add_solar(system_kw = input$system_kw) |>
    dispatch_battery(
      capacity_kwh = input$battery_kwh,
      round_trip_efficiency = input$battery_efficiency
    )
})
```

Every plot, headline and table calls `scenario_data()`.

## The dependency graph is the design

::: {.columns}
::: {.column width="30%"}
### Inputs

- date range
- system size
- tariffs
- battery assumptions
:::

::: {.column width="35%"}
### Shared model

- solar production
- battery dispatch
- grid import/export
- costs and savings
:::

::: {.column width="35%"}
### Views

- salesperson
- neutral analyst
- sceptic
- downloadable evidence
:::
:::

::: {.callout-tip .fragment}
When a number differs across views, the first place to look is the dependency graph.
:::

## Derive the claims from the model

```r
metrics <- reactive({
  df <- scenario_data()

  tibble(
    grid_import_kwh = sum(df$grid_import_kwh),
    grid_export_kwh = sum(df$grid_export_kwh),
    bill_without_system = sum(df$energy_kwh) * input$import_tariff,
    bill_with_system = sum(df$grid_import_kwh) * input$import_tariff -
      sum(df$grid_export_kwh) * input$feed_in_tariff
  ) |>
    mutate(saving = bill_without_system - bill_with_system)
})
```

The narrative is calculated, not typed around a screenshot.

# Turn calculations into three views

## Tabs separate questions—not datasets

```r
tabsetPanel(
  id = "frame",
  tabPanel("Why invest", plotOutput("sales_plot"),
           textOutput("sales_message")),
  tabPanel("Energy balance", plotOutput("neutral_plot"),
           textOutput("neutral_message")),
  tabPanel("Cautious case", plotOutput("sceptic_plot"),
           textOutput("sceptic_message"))
)
```

Each label names the question or perspective the user will encounter.

## Reactive text makes the framing explicit

```r
output$sales_message <- renderText({
  glue::glue(
    "The selected system avoids {scales::dollar(metrics()$saving)} ",
    "in electricity costs over this period."
  )
})

output$sceptic_message <- renderText({
  glue::glue(
    "Simple payback is approximately ",
    "{round(metrics()$payback, 1)} years."
  )
})
```

The code selects a consequence; it does not invent a result.

## The salesperson foregrounds avoided cost

### Question: “What do I gain?”

- lead with cumulative bill savings;
- show how the battery redirects surplus production;
- use confident, specific language; and
- keep costs and assumptions accessible.

::: {.callout-note .fragment}
Persuasion is not permission to hide the feed-in tariff, capital cost or model limitations.
:::

## The neutral view explains the mechanism

### Question: “What is happening to the energy?”

- show consumption, solar production and grid imports together;
- report imports, exports and self-sufficiency;
- state which quantities are observed and modelled; and
- make the selected period and tariffs visible.

This is the reference view against which the other framings can be checked.

## The sceptic keeps the opportunity cost visible

### Question: “What could make this unattractive?”

- subtract up-front costs from cumulative savings;
- highlight grid dependence during low-production periods;
- show the simple payback implied by the same assumptions; and
- distinguish uncertainty from evidence against the system.

“Cautious” should mean conservative interpretation—not a different dataset.

## Three headlines, one model

::: {.columns}
::: {.column width="33%"}
### Why invest

“Save **$X** over the selected period.”
:::

::: {.column width="33%"}
### Energy balance

“Supply **Y%** of consumption without grid imports.”
:::

::: {.column width="33%"}
### Cautious case

“Simple payback is approximately **Z years**.”
:::
:::

::: {.fragment}
The numbers **X, Y and Z** come from the same reactive model.
:::

# Model the battery honestly

## Surplus is not automatically a battery saving {.smaller}

Without a battery model, we can calculate only:

> surplus solar generation that is not consumed at the time it is produced.

Battery savings also require:

- storage capacity;
- charging and discharging losses;
- state of charge; and
- remaining grid imports and exports.

## Conditional controls reveal relevant assumptions

```r
checkboxInput("include_battery", "Include a battery", TRUE)

conditionalPanel(
  condition = "input.include_battery",
  numericInput("battery_kwh", "Usable battery capacity (kWh)",
               value = 10, min = 1, max = 30),
  sliderInput("battery_efficiency", "Round-trip efficiency",
              min = 0.70, max = 1, value = 0.90)
)
```

Hide irrelevant controls, not inconvenient evidence.

# Design for incomplete and awkward states

## A blank output is not an explanation

```r
scenario_data <- reactive({
  req(input$date_range)

  df <- consumption_data |>
    filter(date >= input$date_range[1],
           date <= input$date_range[2])

  validate(need(nrow(df) > 0,
                "No records match this period. Try a wider date range."))

  build_scenario(df, input)
})
```

Validation returns control to the user.

## Every state needs a useful next action {.smaller}

| State | Useful message |
|---|---|
| Waiting | Choose a period and system size. |
| Empty | No records match. Try a wider range. |
| Invalid | Battery capacity must exceed zero. |
| Failed | Check the input data format and try again. |

## Download the evidence, not merely the picture

```r
output$download_data <- downloadHandler(
  filename = function() paste0("solar-scenario-", Sys.Date(), ".csv"),
  content = function(file) {
    readr::write_csv(scenario_data(), file)
  }
)
```

A reader should be able to inspect the common calculation behind every frame.

# Demo and discussion

## Watch what changes—and what does not

During the demo, track four things:

1. Which inputs invalidate the shared model?
2. Which numbers remain identical across tabs?
3. Which design choices make each account feel persuasive?
4. Where could emphasis become misleading?

## Exit ticket

Choose one difference between the salesperson and sceptical views.

Complete both sentences:

> This is a legitimate framing choice because…

> It would become misleading if…

## Week 10 summary

- One shared reactive model keeps the evidence consistent.
- Tabs organise competing questions, not competing calculations.
- Reactive text turns computed results into explicit claims.
- Conditional UI reveals assumptions when they become relevant.
- Helpful states and downloadable evidence make persuasion inspectable.

::: {.callout-important .fragment}
The dashboard’s argument may change. Its underlying evidence must not.
:::
