---
pagetitle: "ETC5523: Communicating with Data"
subtitle: "Stylishly communicating with code"
author: "Michael Lydeamore"
email: "michael.lydeamore@monash.edu"
date: "Week 10"
department: "Department of Econometrics and Business Statistics"
unit-url: "cwd.numbat.space"
footer: "ETC5523 Week 10"
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: c
    multiplex: false
    theme: ../assets/monash.scss
    show-slide-number: all
    show-notes: true
    controls: true
    width: 1280
    height: 720
    css: [../assets/tachyons-addon.css, ../assets/custom.css]
    include-after-body: "../assets/after-body.html"
    incremental: true
execute:
  echo: true
  eval: true
---

```{r, include = FALSE}
current_file <- knitr::current_input()
basename <- gsub(".[Rq]md$", "", current_file)

knitr::opts_chunk$set(
  fig.align = "center",
  out.width = "100%",
  fig.retina = 3,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  cache.path = "cache/"
)
```

## <br>[`r rmarkdown::metadata$pagetitle`]{.monash-blue} {#etc5523-title background-image="images/bg-01.png"}

### `r rmarkdown::metadata$subtitle`

Lecturer: *`r rmarkdown::metadata$author`*

`r rmarkdown::metadata$department`

::: tl
<br>

<ul class="fa-ul">

<li>

[<i class="fas fa-envelope"></i>]{.fa-li}`r rmarkdown::metadata$email`

</li>

<li>

[<i class="fas fa-calendar-alt"></i>]{.fa-li} `r rmarkdown::metadata$date`

</li>

<li>

[<i class="fa-solid fa-globe"></i>]{.fa-li}<a href="`r rmarkdown::metadata[["unit-url"]]`">`r rmarkdown::metadata[["unit-url"]]`</a>

</li>

</ul>

<br>
:::


## {#aim}

::: {.callout-important }

## Aim

* Document your functions
* Create vignettes for your package
* Distribute your package

:::

::: {.callout-tip }

## Why

* Documentation informs users of how to use your package 
* Adopting best practice development workflow will make package development easier 
* Distributing your package is needed for adoption of your package by others


:::

# Recap of package development

## Communicating about your [R package]{.monash-blue}

- What is the [**goal**]{.monash-blue} of the package?
- [**What**]{.monash-blue} does your function(s) do? 
- [**How**]{.monash-blue} do we use it?
- [**Why**]{.monash-blue} should we use it?
- [**Where**]{.monash-blue} do we find and install it?

. . . 

::: {.center .f-headline}

[**Documentation**]{.monash-ruby2} is vital 

:::

## `random_number()` function

```{r}
random_number <- function(range) {
  random_numbers <- data.frame(numbers = runif(n = 100, min = 0, max = range))

  ggplot2::ggplot(data = random_numbers, ggplot2::aes(x = numbers)) +
    ggplot2::geom_histogram()
}

random_number(50)
```

## Using our own data

`data-raw/census-birthplace.R`:
```{r}
#| eval: false
census_birthplace <- read.csv("data-raw/census-birthplace.csv") |>
  dplyr::mutate(birth = factor(birth))

usethis::use_data(census_birthplace, overwrite = TRUE)
```

::: {.fragment}
Then in a package function: 
```{r}
#| eval: false

top_n_birthplace <- function(n, years = c(2016, 2021)) {
  census_birthplace |>
    dplyr::filter(census %in% years) |>
    dplyr::group_by(census) |>
    dplyr::arrange(count) |>
    dplyr::slice(1:n)
}

```
:::


## Using pipe operator in your package 

The base pipe (`|>`) is fine for 99% of cases - I recommend just using that.

If you _need_ the `magrittr` pipe (`%>%`) then you can use

```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_pipe()
```

## Including Shiny apps in your package {.nonincremental}

The process to include a Shiny app is a bit more complicated than data or a function.

Put your app code in `inst/[appname]/app.R`

::: {.fragment}
In your function to launch your app:    
```{r}
#| eval: false

run_app <- function() {
  app_dir <- system.file("ausbirthapp", package = "ausbirthplace")
  shiny::runApp(app_dir, display.mode = "normal")
}

```

:::

::: {.fragment}
Files in `inst/` are installed with your package but are considered "internal".

The `system.file` tells R where to look for your shiny app code.
:::

## Including Shiny apps in your package

::: {.callout-warning}
Don't include an example for running a shiny app! It will cause R to hang.
:::

# Demo


# Documentation {background-color="#006DAE"}


## Documenting R functions with `roxygen2`

- use `#'` above a function to write documentation for that function
- `roxygen2` uses `@` tags to structure documentation, e.g. 
  * any text after `@description` is the description
  * any text after `@param` describes the arguments of the function
  * `@export` signals that it is an exported function
  * any text after `@return` describes the return object
  * the full list of Rd tags are found [here](https://roxygen2.r-lib.org/articles/rd.html)
- `devtools::document()` converts the Rd tags to appropriate sections of `.Rd` files written in the `man/` folder

## Documenting `praise.me` package 

`R/random_number.R`
```{r}
#| eval: false
#' Histograms of random numbers
#'
#' Generates 100 uniform random numbers between 0 and `range`, and
#' plots a `ggplot2` histogram.
#'
#' You can modify the plot using
#' ```r
#' random_number(5) + ggplot2::theme_bw()
#' ```
#'
#' @param range **Maximum** value for the uniform distribution
#'
#' @return `ggplot2` histogram of the random values
#' @export
#'
#' @importFrom ggplot2 ggplot aes geom_histogram
#'
#' @examples
#' random_number(12)
random_number <- function(range) {
  random_numbers <- data.frame(numbers = runif(n = 100, min = 0, max = range-1))

  ggplot2::ggplot(data = random_numbers, aes(x = numbers)) +
    geom_histogram()
}
```


## Documenting data

- `usethis::use_data_raw()` to store R code to process raw data, 
- `usethis::use_data()` to save a binary file in `data/` directory,
- The data is named `praises`.
- Documentation is contained in `data.R` or `name-of-data.R`

## Documenting data

`R/data.R`

```{r}
#| eval: false
#' Number of Australian residents by country of birth from the 2016 and 2021 census
#'
#' @format ## 'census_birthplace'
#' A data frame with 105 rows and 4 columns
#' \describe{
#'   \item{birth}{Country of birth}
#'   \item{count}{Number of residents from that birth country}
#'   \item{percentage}{Percentage of residents from that birth country}
#'   \item{census}{Year of census (either 2016 or 2021)}
#' }
#' @source <https://www.abs.gov.au/websitedbs/censushome.nsf/home/2016>
"census_birthplace"
```

## Make package documentation

- Add documentation of the "big picture" of your package
 
```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_package_doc()
```

- Above creates the file below

. . . 

`R/ausbirthplace-package.R`
```{r}
#| eval: false
#' @keywords internal
"_PACKAGE"

## usethis namespace: start
## usethis namespace: end
NULL
```


- Default package documentation is built from your DESCRIPTION file

. . . 


```{r}
#| eval: false
#| code-line-numbers: false
library(ausbirthplace)
?ausbirthplace
```



## Vignette: a long-form documentation

- Some documentation doesn't fit as a package or function documentation.
- You may want to built a vignette (article) for these cases.

. . . 

```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_vignette(name = "my-amazing-package", 
                      title = "My amazing package")
```

- Edit the created Rmd file 
- Knit the vignette to see what it looks like
- Use `devtools::build()` to build package with vignettes included



# Dependencies {background-color="#006DAE"}

## Adding dependencies


- Dependencies are specified in DESCRIPTION file under three categories:
  - `Depends`: Specify the version of R that the package will work with or package that it is dependent on (e.g. for ggplot2 extension packages, it depends on ggplot2).
  - `Imports`: External packages that are imported to use in your package. Most external packages are in this category.
  - `Suggests`: Packages that are not strictly needed but are nice to have, i.e. you use them in examples or vignettes. 
- You can add easily add this via `usethis::use_package()`

# Sharing {background-color="#006DAE"}

## Share and collaborate on your package


* Track changes to your code with Git

. . . 

```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_git()
```

* Collaborate with others via GitHub (or otherwise)

. . . 

```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_github()
```

. . . 

or for existing repo, run from the terminal:

```{zsh}
#| eval: false
#| code-code-line-numbers: false
git remote add origin https://github.com/user/repo.git
```


* You can install your R package now using:

. . . 


```{r}
#| eval: false
#| code-line-numbers: false
devtools::install_github("user/repo")
```

## Installing `praise.me` package

```{r}
#| eval: false
#| code-line-numbers: false
devtools::install_github("MikeLydeamore/ausbirthplace")
```

* The package is found at `https://github.com/MikeLydeamore/ausbirthplace`.
* It's a good idea to add a `README` file with installation instructions -- this is displayed in the GitHub repo.
* You can create a `README.Rmd` file with

. . . 

```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_readme_rmd() 
# OR usethis::use_readme_md() if you have no code
```
* Make sure you knit the `README.Rmd` when you modify its contents.

## Package documentation website with `pkgdown`

- Automatically turns all package documentation into a website.
- Documentation can now be easily viewable outside of R.
- Easy to customise appearance of the site using YAML

## Using `pkgdown` 


```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_pkgdown()
```

- Build site locally with `pkgdown::build_site()`
- Site appearance is modified in the `_pkgdown.yml` file
  - [bootswatch](https://bootswatch.com/) themes for the appearance of the whole site
  - organising function / vignette documentation with reference
- See the [vignette](https://pkgdown.r-lib.org/articles/pkgdown.html) for more details
- Automatically build and deploy your site with GitHub actions

. . . 

```{r}
#| eval: false
#| code-line-numbers: false
usethis::use_pkgdown_github_pages() # if using this, no need for usethis::use_pkgdown()
```

## The whole package development workflow

```{r}
#| eval: false
available::available("pkgname") # check if package name is available (if planning to publish publicly)
usethis::create_package("pkgname")
usethis::use_git() # set up version control
usethis::use_github() # optional
usethis::use_r("myfile")
# write some functions in a script
usethis::use_data_raw() # if adding data
devtools::load_all() # try it out in the console
usethis::use_package("import-pkgname") # add package to import (or depends or suggests)
usethis::use_package_doc() # add package documentation
usethis::use_pipe() # if you want to add %>% from `magrittr`
usethis::use_vignette("vignette-name") # add vignette
devtools::build() # build vignettes
devtools::install() # to install package
devtools::check() # to build and check a package 
usethis::use_readme_rmd() # to add a README Rmd file
styler::style_pkg() # optional (commit first, then review changes carefully)
usethis::use_pkgdown_github_pages() # for setting up pkgdown website on github
# `usethis::use_pkgdown()` if not using github pages
```




## Week 10 Lesson 


::: callout-important

## Summary

::: nonincremental

- Package documentation is important to let others know about the goal of the package, what your function does, and how to use your package.
- Sharing your package by making it easy to install, commiting to good documentation, and making the documentation accessible helps to build trust to use your package.
- You can make package development and distribution easy with `usethis`, `devtools`, `roxygen2`, and `pkgdown`.

:::

:::

. . . 

::: callout-tip

## Resources

::: nonincremental


- [`roxygen2` documentation tags](https://roxygen2.r-lib.org/articles/rd.html)
- [Customising your `pkgdown` site](https://pkgdown.r-lib.org/articles/customise.html)

:::

:::
