---
title: 'ETC5523: Communicating with Data'
subtitle: "Assignment 3: Creating a website"
author: "Michael Lydeamore"
date: today
format:
    html:
        embed-resources: true
---

```{r}
#| echo: false

total_marks <- 0

add_marks <- function(amount) {
    total_marks <<- total_marks + amount

    amount
}

```

```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```

## Objectives

- Create your own personal website
- Adapt your Assignment 2 report into a blog post for a clearly defined audience
- Explain and justify your communication and visual-design decisions
- Communicate your data story clearly in a short recorded walkthrough

::: callout-important
This is an **individual assignment**.
:::

## Output format

- A published website, using Quarto
- A recorded screen-and-voice walkthrough of the published website

## Instructions

In this assessment, you will create a personal website, including information about you, and adapt your Assignment 2 report into a blog post for a clearly defined audience.

::: {.callout-important}

You must join the Classroom50 Assignment here: [https://classroom50.org/ETC5523-2026/etc5523/assignments/assignment-3/accept](https://classroom50.org/ETC5523-2026/etc5523/assignments/assignment-3/accept). **If you cannot find your name in the list of students, please contact the lecturer directly before accepting the assignment.**

:::

### Part A: The homepage (3 marks)

The template provided in GitHub classroom will set you up a basic website. Edit the `_quarto.yml` file to include:

* A sensible title for your website and a link to _your_ GitHub account (1 mark)
* Some custom theming (2 marks):
    * Theming is subjective, but make sure to consider readability of your website. As an example, changing the default background colours sensibly could give 1 mark, and changing font colours for headings another mark.
    * You are encouraged to seek feedback on your theming before submission.

::: {.callout-tip}
The default theming is deliberately poor!
:::

The listing page will automatically generate, and you do not need to make any changes here.

### Part B: The about page (2 marks)

In the about page (`about.qmd`), make the following changes:

* Add a photo (or some representation) of you and write a short biography (1 mark)
* Change the GitHub button to link to your GitHub, and add any other buttons of your choosing (1 mark)

For an example about page, you can see mine [here](https://www.michaellydeamore.com/)

### Part C: Blog post

There is an example blog post included in the template. Replace this post with one about the story that you have written for Assignment 2.

::: {.callout-important}
Don't just copy-paste Assignment 2 into your blog post. Blog posts use different language and features to appeal to a different audience.
:::

Choose a clearly defined intended audience for your blog post. Add a one-sentence audience statement to the repository README so that the audience is clear to the marker. Your choices about language, information order, figures and tables should be appropriate for this audience.

The final blog post must have:

* A title and publishing date
* Correct author information (you)
* The following sections:
    * Problem description
    * Analysis
    * Conclusion
* At least one figure, with a cross-reference in text
* At least one table, with a cross-reference in text
* At least one academic citation

You don't have to use these headings.

#### Marking rubric for Part C

Problem description (2 marks):

* Short summary of the problem or issue, written at an appropriate level for the intended audience (2 marks)

Analysis (10 marks):

* Source of the evidence or data used in the original article clearly identified (1 mark)
* Findings reported in the original article clearly described in language appropriate for the intended audience (2 marks)
* Tables and plots designed so that the intended audience can acquire the desired information efficiently (2 marks for plot, 2 marks for table)
* Included cross-references (1 mark)
* Discussion of the evidence links back to the questions in the problem description (2 marks)


Conclusion (5 marks):

* Links back to the problem description (2 marks)
* Includes academic citation (here or other section) (1 mark)
* Summarises the findings (2 marks)

### Part D: Recorded website walkthrough (8 marks)

Record an individual **4--6 minute** screen-and-voice walkthrough of your published website. The published website must be visible as you navigate it and discuss your work. The recording should be a single continuous take, led by your spoken explanation rather than by text on screen. Do not submit a slide deck or a narrated set of slides in place of the website walkthrough.

Upload this recording as your assignment submission on Moodle as a video file.

Your walkthrough must:

* identify the intended audience and introduce the problem or question;
* explain the most important finding(s), referring to visual evidence on the published website;
* identify and explain **two specific changes** you made when adapting your Assignment 2 report for the blog's intended audience;
* explain one design choice you made for a figure or table;
* identify one important limitation or qualification that the audience should understand; and
* end with a clear conclusion or implication for the audience.

#### Marking rubric for Part D

* Clear, accurate explanation of the data story and its main finding(s) (2 marks)
* Explanation of two specific adaptations and why they suit the intended audience (2 marks)
* Explanation of one figure or table design choice (1 mark)
* Clear explanation of an important limitation or qualification (1 mark)
* Logical navigation and clear, concise verbal delivery within the time limit (2 marks)

### Part E: Technical work

Publish your website using _GitHub pages_.

The website must be publicly accessible from a URL - the default URL is perfectly fine.

To do this, you will need to:

* Set up Pages on your Classroom repository
* Change the visibility of your repository to _Public_
* Publish your website using either `quarto publish` or by having the site directory as part of your repo

#### Marking rubric for Part E

* Working link (1 mark)
* Pages set up correctly, with the website HTML files correctly under version control (1 mark)

**Maximum grade**: 3 + 2 + 17 + 8 + 2 = 32 marks
