---
title: 'ETC5523: Communicating with Data'
subtitle: "Assignment 2: Summarising technical content"
author: "Michael Lydeamore"
date: today
format:
    html:
        embed-resources: true
---

```{r}
#| echo: false

total_marks <- 0

add_marks <- function(amount) {
    total_marks <<- total_marks + amount

    amount
}

```

```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```

## `r emo::ji("target")` Objectives

-   Create a piece of technical communication

::: callout-important
This is an **individual assignment**.
:::

## Output format

-   Maximum length: Four A4 pages (Part A) + 2 A4 pages (Part B)

-   Please upload the completed PDF to Moodle

## Instructions

In this asseessment, you will be transforming a piece of academic writing into an executive-focussed technical report. You can re-use any tables and figures from the original article, but you will need to write your own text.

### Part A: Writing the article

You are an analyst in a research company. You have noticed some new academic work that is relevant to your field. The first step to bringing the work in house is to write a technical report for your manager.

Your manager is competent, and reasonably technical. They are time-poor, and so clear communication will help make your point the fastest. Your report will need to include:

* An executive summary
* An introduction to the problem
* Key results
* **Brief** methods
* Conclusion

Your report also needs:

* At least one reference to an article (academic, web page etc)
* Working cross-references to figure(s), table(s) if included, and sections where required

::: {.callout-caution}
Interpreting these papers **is difficult**. You are not expected to understand all of the technical details. If you have _any_ questions about the papers, please ask in the forums, or during consultation. There will be a couple of dedicated question times in the workshops on weeks with less oral presentations.
:::

#### Choosing an article

Four articles are available:

* [COVID-19 in low-tolerance border quarantine systems: Impact of the Delta variant of SARS-CoV-2](assignment-2-papers/Zachreson2021.pdf)
* [Application of a new methodology and R package reveals a high burden of healthcare-associated infections (HAI) in Germany compared to the average in the European Union/European Economic Area, 2011 to 2012](assignment-2-papers/Zacher2019.pdf)
* [Attribution of the Australian bushfire risk to anthropogenic climate change](assignment-2-papers/Oldebborgh2021.pdf)
* [Understanding links between water-quality variables and nitrate concentration in freshwater streams using high frequency sensor data](assignment-2-papers/Kermorvant2023.pdf)

You may also choose your own article not in this list, but it must be approved by the CE.

### Part B: Article analysis

Compare and contrast the format of the article you chose in Part A with the report you wrote. In particular, consider:

* Where the key takeaways are located
* How results are presented
* Caveats and limitations in both the article and your report

## Marking Rubric

Part A (15 marks):

- Structure adhered to (1 mark)
- Executive summary summarises the work (3 marks)
- Report is clear and easy to understand (3 marks)
- Report is written at the correct level (1 mark)
- Paragraphs contain topic sentences, and flow from one to the next (2 marks)
- Results include a figure and/or a table, with a clear caption, and a cross-reference in text (2 marks)
- Methods includes data source, and some details of work done (2 marks)
- References included (1 mark)

Part B (5 marks): 

-   Compared and contrasted differences
-   Highlighted reasons for differences

**Maximum grade:** 20 marks

Up to 3 marks will be deducted for spelling and grammatical errors.