---
title: 'ETC5523: Communicating with Data'
subtitle: "Assignment 3: Creating a website"
author: "Michael Lydeamore"
date: today
format:
    html:
        embed-resources: true
---

```{r}
#| echo: false

total_marks <- 0

add_marks <- function(amount) {
    total_marks <<- total_marks + amount

    amount
}

```

```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```

## `r emo::ji("target")` Objectives

- Create your own personal website
- Turn your previous article into a blog post

::: callout-important
This is an **individual assignment**.
:::

## Output format

- Published website, using Quarto

## Instructions

In this assessment, you will be creating a personal website, including information about you, and a blog post for the story you have worked on in Assignment 2.

::: {.callout-important}

You must join the GitHub classroom here: [https://classroom.github.com/a/pM7OZP_9](https://classroom.github.com/a/pM7OZP_9). **If you cannot find your name in the list of students, please contact the lecturer directly before accepting the assignment.**

:::

### Part A: The homepage (4 marks)

The template provided in GitHub classroom will set you up a basic website. Edit the `_quarto.yml` file to include:

* A sensible title for your website (1 mark)
* A link to _your_ GitHub account (1 mark)
* Some custom theming (2 marks):
    * Theming is subjective, but make sure to consider readability of your website. As an example, changing the default background colours sensibly could give 1 mark, and changing font colours for headings another mark.
    * You are encouraged to seek feedback on your theming before submission.

::: {.callout-tip}
The default theming is deliberately poor!
:::

The listing page will automatically generate, and you do not need to make any changes here.

### Part B: The about page (3 marks)

In the about page (`about.qmd`), make the following changes:

* Add a photo of you
* Write a short biography
* Change the GitHub button to link to your GitHub, and add any other buttons of your choosing

For an example about page, you can see mine [here](https://www.michaellydeamore.com/)

### Part C: Blog post

There is an example blog post included in the template. Replace this post with one about the story that you have written for Assignment 2.

::: {.callout-important}
Don't just copy-paste Assignment 2 into your blog post. Blog posts use different language, and have different features to appeal to a difference audience.
:::

The final blog post must have:

* A title and publishing date
* Correct author information (you)
* The following sections:
    * Problem description
    * Data description
    * Analysis
    * Conclusion
* At least one figure, with a cross-reference in text
* At least one table, with a cross-reference in text
* At least one academic citation

You don't have to use these headings.

#### Marking rubric for Part C

Problem description (2 marks):

* Short summary of the problem or issue you are analysing (2 marks)

Data description (6 marks):

* Source of the data clearly identified (1 mark)
* Short description of the variables in the dataset (not a data dictionary!) (2 marks)
* Data cleaning steps described (2 marks)
* Code included where needed (which may not be everywhere) (1 mark)

Analysis (5 marks):

* Tables and plots designed so that the audience can acquire the desired information efficiently (2 marks for plot, 2 marks for table)
* Included cross-references (1 mark)
* Analysis links back to the questions in the problem description (2 marks)


Conclusion (5 marks):

* Links back to the problem description (2 marks)
* Includes academic citation (here or other section) (1 mark)
* Summarises the findings (2 marks)

### Part D: Technical work

Publish your website using _GitHub pages_.

The website must be publicly accessible from a URL - the default URL is perfectly fine.

To do this, you will need to:

* Set up Pages on your Classroom repository
* Change the visibility of your repository to _Public_
* Publish your website using either `quarto publish` or by having the site directory as part of your repo

#### Marking rubric for Part D

* Working link (1 mark)
* Pages set up correctly (1 mark)
* Website HTML files correctly on version control (1 mark)

**Maximum grade**: 4 + 3 + 18 + 3 = 28 marks