---
title: 'ETC5523: Communicating with Data'
subtitle: "Oral Presentations"
author: "Michael Lydeamore"
date: today
format:
    html:
        embed-resources: true
---

```{r}
#| echo: false

total_marks <- 0

add_marks <- function(amount) {
    total_marks <<- total_marks + amount

    amount
}

```

```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```

## `r emo::ji("target")` Objectives

-   Practice your verbal communication skills to your peers

::: callout-important
This is an **individual** assignment.

The final PDF report can not have more than 5 pages (not including the appendix).

You cannot use a dataset that has been used for another assignment.
:::

## Topics

| Week | Topic 1                                              | Topic 2                                         | Topic 3                                                               |
|---------------|------------------|----------------|------------------------|
| 2    | Describe the art of rhetorics                        | Demonstrate how to ask a good question, and why | The differences between linear and transactional communication models |
| 3    |                                                      |                                                 |                                                                       |
| 4    | Different narrative structures for different reports | Richard Mayer's learning principles             |                                                                       |
| 5    | Quarto vs RMarkdown: Key differences                 | Writing Markdown for reports and slides         | Web directories and servers                                           |
| 6    | Which statistics should you present?                 | Key elements of a good table                    |                                                                       |
| 7    |                                                      |                                                 |                                                                       |
| 8    |                                                      |                                                 |                                                                       |
| 9    |                                                      |                                                 |                                                                       |
| 10   |                                                      |                                                 |                                                                       |
| 11   |                                                      |                                                 |                                                                       |
| 12   |                                                      |                                                 |                                                                       |