---
title: 'ETC5523: Communicating with Data'
subtitle: "Oral Presentations"
author: "Michael Lydeamore"
date: today
format:
    html:
        embed-resources: true
---

```{r}
#| echo: false

total_marks <- 0

add_marks <- function(amount) {
    total_marks <<- total_marks + amount

    amount
}

```

```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```

## `r emo::ji("target")` Objectives

-   Practice your verbal communication skills to your peers

::: callout-important
In Week 1, your tutor will have a sign-up sheet for presentations. **Please only sign up to a tutorial you can attend!**

If you miss your tutorial, then you can either:

-   Pick another topic that hasn't already been signed up for or,

-   Schedule your presentation to be performed within 7 days with the Chief Examiner. Please email the [Chief Examiner](mailto:michael.lydeamore@monash.edu) to set up this time.
:::

## Presentation format

-   Maximum time: 5 minutes

-   **No slides**

-   Concise and clear communication will be rewarded.

## Topics

| Week | Topic 1                                              | Topic 2                                         | Topic 3                                                               |
|-----------------|-----------------|-----------------|----------------------|
| 2    | Describe the art of rhetorics                        | Demonstrate how to ask a good question, and why | The differences between linear and transactional communication models |
| 3    |                                                      |                                                 |                                                                       |
| 4    | Different narrative structures for different reports | Richard Mayer's learning principles             |                                                                       |
| 5    | Quarto vs RMarkdown: Key differences                 | Writing Markdown for reports and slides         | Web directories and servers                                           |
| 6    | Which statistics should you present?                 | Key elements of a good table                    |                                                                       |
| 7    |                                                      |                                                 |                                                                       |
| 8    |                                                      |                                                 |                                                                       |
| 9    |                                                      |                                                 |                                                                       |
| 10   |                                                      |                                                 |                                                                       |
| 11   |                                                      |                                                 |                                                                       |
| 12   |                                                      |                                                 |                                                                       |

## Marking Rubric

**Maximum grade:** 5 marks

0.  No presentation, or completely incorrect information
1.  Presentation has clearly incorrect elements with little to no understanding
2.  Presentation shows limited understanding, insufficient detail provided
3.  Presentation shows some understanding, with some detail
4.  Presentation gives good accompanying details, and a strong understanding of the topic
5.  Presentation is complete and comprehensive without being verbose. Demonstrated clear understanding of the topic and has explained the details clearly.

1 mark deduced per 30 seconds over 5 minutes.