---
title: 'ETC5523: Communicating with Data'
subtitle: "Oral Presentations"
author: "Michael Lydeamore"
date: today
format:
    html:
        embed-resources: true
---

```{r}
#| echo: false

total_marks <- 0

add_marks <- function(amount) {
    total_marks <<- total_marks + amount

    amount
}

```

```{r, echo = FALSE, eval = file.exists("tutorial.css")}
htmltools::includeCSS("tutorial.css")
```

::: {.callout-important}
You can view the oral presentation schedule online [here](https://docs.google.com/spreadsheets/d/1xfcRph4U4IBxIv0ZNAoro_zlHLxVpwOLwlMghHSmK9w/edit?usp=sharing).
:::

## `r emo::ji("target")` Objectives

-   Practice your verbal communication skills to your peers

::: callout-important

Please sign up to an oral presentation spot here: [Oral Presentation Sign Up](https://docs.google.com/spreadsheets/d/1UK107GEnyjJSAJKKy2E41VcFVyMlSxz2XbUQ3V85Jnk/edit?usp=sharing).
:::

This assignment makes up 5% of your final grade for the unit.

## Presentation format

-   Maximum time: 5 minutes
-   **No slides**
-   Concise and clear communication will be rewarded.

## Topics

| Week | Topic 1                                              | Topic 2                                          | Topic 3                                                               |
|------------------|------------------|------------------|--------------------|
| 2    | The art of rhetorics                                 | Demonstrate how to ask a good question, and why  | The differences between linear and transactional communication models |
| 3    | Best practices when including data in your narrative | What are the elements of a good story?           | How to capture an audience                                            |
| 4    | Different narrative structures for different reports | Richard Mayer's learning principles              | Key elements for presenting data in a presentation                    |
| 5    | Quarto vs RMarkdown: Key differences                 | Writing Markdown for reports and slides          | Web directories and servers                                           |
| 6    | Which statistics should you present?                 | Key elements of a good table                     |                                                                       |
| 7    | Choices of plots for different purposes              | Importance of unifying visual elements in a plot | Different graphics libraries for producing plots                      |
| 8    | How are web pages styled?                            | Writing in HTML vs PDF vs docx                   | Benefits of using Quarto for scientific writing                       |
| 9    | Basic elements of good function documentation        | Key roxygen elements and their purpose           | R Package structure                                                   |
| 10   | Why should we use a dashboard?                       | Reactive programming and shiny                   | Deploying web applications                                            |
| 11   | Communicating with code: Key concepts                | Workflows for deploying package documentation    | How is code documentation part of communication?                      |
| 12   | How do we ensure we can find information quickly?    | Challenges of data governance when communicating | What are our responsibilities when publishing mass communication?     |

## Marking Rubric

**Maximum grade:** 5 marks

0.  No presentation, or completely incorrect information
1.  Presentation has clearly incorrect elements with little to no understanding
2.  Presentation shows limited understanding, insufficient detail provided
3.  Presentation shows some understanding, with some detail
4.  Presentation gives good accompanying details, and a strong understanding of the topic
5.  Presentation is complete and comprehensive without being verbose. Demonstrated clear understanding of the topic and has explained the details clearly.

1 mark deduced per 30 seconds over 5 minutes. A 5 minute and 1 second presentation will receive a maximum of 4 marks.