---
pagetitle: "ETC5523: Communicating with Data"
subtitle: "R Packages and documentation"
author: "Michael Lydeamore"
email: "michael.lydeamore@monash.edu"
date: "Week 8"
department: "Department of Econometrics and Business Statistics"
unit-url: "cwd.numbat.space"
footer: "ETC5523 Week 8"
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: c
    multiplex: false
    theme: ../assets/monash.scss
    show-slide-number: all
    show-notes: true
    controls: true
    width: 1280
    height: 720
    css: [../assets/tachyons-addon.css, ../assets/custom.css]
    include-after-body: "../assets/after-body.html"
    incremental: false
execute:
  echo: true
  eval: true
---

```{r, include = FALSE}
current_file <- knitr::current_input()
basename <- gsub(".[Rq]md$", "", current_file)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 3,
  warning = FALSE,
  message = FALSE,
  cache = FALSE
)
```

## <br>[`r rmarkdown::metadata$pagetitle`]{.monash-blue} {#etc5523-title background-image="images/bg-01.png"}

### `r rmarkdown::metadata$subtitle`

Lecturer: *`r rmarkdown::metadata$author`*

`r rmarkdown::metadata$department`

::: tl
<br>

<ul class="fa-ul">

<li>

[<i class="fas fa-envelope"></i>]{.fa-li}`r rmarkdown::metadata$email`

</li>

<li>

[<i class="fas fa-calendar-alt"></i>]{.fa-li} `r rmarkdown::metadata$date`

</li>

<li>

[<i class="fa-solid fa-globe"></i>]{.fa-li}<a href="`r rmarkdown::metadata[["unit-url"]]`">`r rmarkdown::metadata[["unit-url"]]`</a>

</li>

</ul>

<br>
:::


## {#aim}

::: {.callout-important }

## Aim

* Document your functions
* Document data and longer workflows
* Check that your package works
* Share your package and its documentation

:::

::: {.callout-tip }

## Why

* Reference documentation answers specific questions
* READMEs and vignettes help new users get started
* Package checks catch problems before users encounter them
* GitHub and `pkgdown` make your work installable and discoverable


:::


# Let's make a package

This is a taste of ETC4500/5450 (Advanced R Programming). We will learn the bare minimum required to get to documentation.

## Communicating about your [R package]{.monash-blue}

- What is the [**goal**]{.monash-blue} of the package?
- [**What**]{.monash-blue} does your function(s) do? 
- [**How**]{.monash-blue} do we use it?
- [**Why**]{.monash-blue} should we use it?
- [**Where**]{.monash-blue} do we find and install it?

. . . 

::: {.center .f-headline}

[**Documentation**]{.monash-ruby2} is vital 

:::

## `pokemonvgc` package

Last week we rescued a plot of Pokémon Video Game Championship Elo ratings. We will turn that work into a small package so that someone else can derive the ratings, use the data, and repeat the analysis.

It's never been easier to make a package. The `usethis` package can do all of the heavy lifting for us

::: {.callout-tip}
This means we can easily make packages that might only be just "for us"
:::

::: {.callout-note appearance="minimal"}

<i class="fa-solid fa-circle-exclamation"></i> This package is for teaching demo only

:::

```{r}
#| eval: false

usethis::create_package("pokemonvgc")

```

will set up the directory structure we need for a package.

## Package directory structure

```
- R/
- data/
- data-raw/
- man/
- vignettes/
- DESCRIPTION
- NAMESPACE
- README.md
```

::: {.callout-tip}
Different forms of communication live in different parts of the package.
:::

## Functions for the package

Functions in the `R/` folder become part of the package.

For `pokemonvgc`, we will export `recent_players()`. It reduces rating histories to the most recent observation for each player with enough recorded events.

::: {.callout-tip}
Add `@export` to functions that users should call directly.
:::

## Data for the package

Our package will include two user-facing data objects:

* `pokemon_elo`: Elo histories derived from the match history
* `recent_champions`: the champion lookup used last week

User-facing data objects go in `data`; the scripts that produce them go in `data-raw`.

The match history is a raw input to the build process, not a dataset shipped
to package users.

The source pipeline downloads and caches VGC History's tournament catalogue and per-event match files, then validates and normalises them.

```{r}
#| eval: false
usethis::use_data_raw("pokemon-elo")
```

## Build the Elo histories

The generated script records how cached match data become package data objects:

```{r}
#| eval: false
source("data-raw/calculate-elo.R")

# Created by the cached VGC History download and validation pipeline
match_history <- readRDS("data-raw/vgc-match-history.rds")
players <- readRDS("data-raw/vgc-players.rds")

ratings <- calculate_elo(match_history) |>
  tibble::as_tibble() |>
  dplyr::group_by(player_id, date) |>
  dplyr::summarise(
    rating = round(dplyr::last(rating), 1),
    .groups = "drop"
  )
```

## Include recent champions

```{r}
#| eval: false
recent_champions <- tibble::tribble(
  ~year, ~player_id,          ~player_name,
  2022,  "eduardo-cunha",    "Eduardo Cunha",
  2023,  "shohei-kimura",    "Shohei Kimura",
  2024,  "luca-ceribelli",   "Luca Ceribelli",
  2025,  "giovanni-cischke", "Giovanni Cischke",
  2026,  "takuma-yamazaki",  "Takuma Yamazaki"
)
```

Keeping this lookup as data separates championship status from the rating calculation.

## Reproduce the workshop snapshot

```{r}
#| eval: false

frequent_players <- ratings |>
  dplyr::count(player_id) |>
  dplyr::filter(n > 35) |>
  dplyr::pull(player_id)

selected_players <- union(frequent_players, recent_champions$player_id)
```

This is the 47 frequently observed players plus the four recent champions represented in the archive.

## Finish and save the package data

```{r}
#| eval: false

pokemon_elo <- ratings |>
  dplyr::filter(player_id %in% selected_players) |>
  dplyr::left_join(
    dplyr::select(players, player_id, player_name),
    by = "player_id"
  ) |>
  dplyr::select(player_id, player_name, date, rating)

usethis::use_data(
  pokemon_elo, recent_champions,
  overwrite = TRUE
)
```

This reproduces the 51 players and 2,043 observations used last week. The 2026 champion remains in the lookup but has no rating history in this snapshot.

## Which code belongs where?

::: {.callout-tip}
Keep downloading, caching, one-off cleaning, and the Elo calculation in `data-raw/`. These steps build the package data but are not part of its user-facing interface.
:::

::: {.callout-note}
Put `recent_players()` in `R/recent-players.R` and export it: selecting the latest eligible players is a task package users will repeat.
:::

## Load the package while developing

`devtools` lets us load the package without installing it first:

```{r}
#| eval: false

devtools::load_all()
pokemon_elo
```

::: {.fragment}
You can also use the default hotkey `Ctrl+Shift+L` to load.
:::

::: {.fragment}
You should reload often - try to avoid manually sourcing files unless you have to!
:::

## NAMESPACE, DESCRIPTION, README

Three files play different communication roles:

::: {.incremental}
* `NAMESPACE`: machine-generated instructions about exported functions and imports
* `DESCRIPTION`: structured metadata about the package and its dependencies
* `README`: the introduction people see _before_ they install the package
:::

## README

[Hadley](https://r-pkgs.org/other-markdown.html#sec-readme) gives a good template for a README:

1. A paragraph describing the purpose of the package
2. An example that shows how to use the package to solve a simple problem
3. Installation instructions that you can copy/paste straight into R
4. An overview of the main components of the package.

::: {.fragment}
We can set up the README using

```{r}
#| eval: false
usethis::use_readme_rmd()
```

and then edit to suit your package.

:::

## README

For `pokemonvgc`, the opening should tell readers that it contains VGC match data, derived Elo histories, and recent champion metadata. Then show the shortest useful example:

```r
library(pokemonvgc)

recent_players(pokemon_elo) |>
  dplyr::slice_max(rating, n = 3)
```

## DESCRIPTION

`DESCRIPTION` tells tools and users what the package is:

```
Package: pokemonvgc
Title: Explore Competitive Pokemon VGC Results
Version: 0.0.0.9000
Authors@R: 
    person("First", "Last", role = c("aut", "cre"))
Description: Provides competitive Pokemon VGC match data, derived Elo
    rating histories, recent champion metadata, and tools for identifying
    each eligible player's latest rating.
License: MIT + file LICENSE
Imports:
    dplyr
```

::: {.callout-important}
Replace every placeholder; a valid package needs a meaningful title and description.
:::

## Choose a licence

A licence tells other people what they may do with your code.

```{r}
#| eval: false

usethis::use_mit_license("Your Name")
```

::: {.callout-tip}
For Assignment 4, include a licence file rather than leaving the generated placeholder in `DESCRIPTION`.
:::

## Using other packages

Inevitably, you'll want to use a function from another package. Again there's a `usethis` function for that too:

```{r}
#| eval: false

usethis::use_package("dplyr")
```

For now, reference exported functions using `::`. After introducing `roxygen2`, we will import selected names where that makes package code easier to read.

For example:

```{r}
#| eval: false

recent_players <- function(ratings, min_events = 15) {
  ratings |>
    dplyr::group_by(player_id, player_name) |>
    dplyr::mutate(events = dplyr::n()) |>
    dplyr::slice_max(date, n = 1, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::filter(events >= min_events)
}
```

## Code commenting {auto-animate="true"}

Comments in code _are communication_. There's no need to re-state very clear code, but if you are doing anything even mildly complicated, chances are you won't be able to understand the code quickly.

Example:

```r
pokemon_elo |>
  dplyr::group_by(player_id, player_name) |>
  dplyr::mutate(events = dplyr::n()) |>
  dplyr::slice_max(date, n = 1, with_ties = FALSE) |>
  dplyr::ungroup() |>
  dplyr::filter(events >= 15) |>
  dplyr::slice_max(rating, n = 3, with_ties = FALSE)
```

## Code commenting {auto-animate="true"}

Comments in code _are communication_. There's no need to re-state very clear code, but if you are doing anything even mildly complicated, chances are you won't be able to understand the code quickly.

Example:

```r
pokemon_elo |>
  dplyr::group_by(player_id, player_name) |>
  # Count events before keeping only each player's latest rating
  dplyr::mutate(events = dplyr::n()) |>
  dplyr::slice_max(date, n = 1, with_ties = FALSE) |>
  dplyr::ungroup() |>
  # Match the workshop definition of an established current player
  dplyr::filter(events >= 15) |>
  dplyr::slice_max(rating, n = 3, with_ties = FALSE)
```

## Code commenting

Useful comments explain intent, assumptions, or a surprising decision.

Avoid narrating code that is already clear. As with all communication, conciseness and clarity are key.

## Function documentation

When we don't know how to use a function, we look up the _function documentation_ by typing `?group_by`.

But where does this documentation come from?

::: {.fragment}
R gives a standard way of documenting packages: you write `.Rd` files in the `man/` directory. These files use a custom syntax, which is very similar to LaTeX.
:::

::: {.fragment}
This separates code from comments, making it easy to forget to update a function's documentation if you change the function. `roxygen2` puts the documentation with your code, and generates `.Rd` files.
:::

## `roxygen2` skeleton

```r
#' Function title
#' 
#' Function description
#' 
#' @param param_name Parameter description
#' @return What does the function return?
#' @examples
#' R code for examples goes here
myfunction <- function(param_name) {
  print(param_name)
}
```

## A function worth documenting {auto-animate="true"}

```r
recent_players <- function(ratings, min_events = 15) {
  ratings |>
    dplyr::group_by(player_id, player_name) |>
    dplyr::mutate(events = dplyr::n()) |>
    dplyr::slice_max(date, n = 1, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::filter(events >= min_events)
}
```

Before reading the source, what would a user need to know?

## Import only the functions you use {auto-animate="true"}

```r
#' @importFrom dplyr filter group_by mutate n slice_max ungroup
recent_players <- function(ratings, min_events = 15) {
  ratings |>
    group_by(player_id, player_name) |>
    mutate(events = n()) |>
    slice_max(date, n = 1, with_ties = FALSE) |>
    ungroup() |>
    filter(events >= min_events)
}
```

`@importFrom` lets package code use those selected names without repeating `dplyr::`.

::: {.callout-tip}
Prefer selective `@importFrom` declarations to importing an entire package with `@import dplyr`.
:::

## Dependency, import, generation

| Instruction | Role |
|---|---|
| `usethis::use_package("dplyr")` | Records `dplyr` in `DESCRIPTION` |
| `@importFrom dplyr ...` | Declares names used inside package code |
| `devtools::document()` | Writes `importFrom(...)` entries to `NAMESPACE` |

::: {.callout-important}
Do not edit `NAMESPACE` by hand. If an imported name is ambiguous, keep the explicit `package::function()` form.
:::

## Describe purpose, inputs, and output

```r
#' Find each eligible player's most recent rating
#' 
#' Keeps the latest rating for every player with at least `min_events`
#' observations in the supplied rating histories.
#'
#' @param ratings A data frame containing `player_id`, `player_name`,
#'   `date`, and `rating`.
#' @param min_events Minimum number of observations a player must have.
#' @return A data frame with one row per eligible player, including their
#'   most recent date and rating and their total number of events.
#' @importFrom dplyr filter group_by mutate n slice_max ungroup
#' @export
recent_players <- function(ratings, min_events = 15) {
  # ...
}
```

## Examples teach use and interpretation

```r
#' @examples
#' recent_players(pokemon_elo)
#' # Each established player's latest available rating
#'
#' recent_players(pokemon_elo, min_events = 10) |>
#'   dplyr::slice_max(rating, n = 3)
#' # The current top three under a lower eligibility threshold
```

::: {.callout-tip}
Examples should be runnable and help users interpret the result.
:::

## Generating the `.Rd` file

Just writing `roxygen2` doesn't generate the man files. To get these, we have to run

`devtools::document()`

::: {.fragment}
If you are in RStudio, you can press ctrl+shift+d
:::

::: {.callout-important .fragment}
For a function to be accessible by installing your package, you need to include the `@export` tag!
:::

## Data documentation

Just like functions, we have to document our data. You can put this anywhere (after all, there is no function that generates our data), but the convention is in `R/data.R`.

For our data,

```r
#' Pokemon Video Game Championship Elo rating histories
#' 
#' @format A data frame with 2,043 rows and 4 columns:
#' \describe{
#'   \item{player_id}{Stable player identifier}
#'   \item{player_name}{Player display name}
#'   \item{date}{Date of the rated event}
#'   \item{rating}{Estimated Elo rating after the event}
#' }
#' @details Elo is estimated from the available match archive and is not
#'   an official ranking. Coverage varies across events.
#' @source Match data compiled by VGC History,
#'   <https://vgchistory.com/data>
"pokemon_elo"
```

Every user-facing data object needs its own documentation block. Raw inputs
kept only in `data-raw/`, such as `match_history`, are documented for package
developers rather than exposed through the package help system.

## Document the champion lookup

```r
#' Recent Pokemon VGC World Champions
#'
#' @format A data frame with 5 rows and 3 columns:
#' \describe{
#'   \item{year}{Championship year}
#'   \item{player_id}{Stable player identifier}
#'   \item{player_name}{Player display name}
#' }
#' @source VGC History, <https://vgchistory.com/>
"recent_champions"
```

::: {.callout-tip}
Metadata can include a player who has no rows in the rating snapshot. Document that distinction rather than silently dropping them.
:::

## Vignettes

Documentation is great to look at something specific (like a function).

What if you've just installed a package and would like to know how to use it?

::: {.fragment}
Enter the Vignette
:::

## Vignettes

Vignettes are "how-to" guides to packages. Perhaps they document a specific workflow, or how to solve a specific problem.

Check out the [dplyr vignettes](https://dplyr.tidyverse.org/articles/)

## Match the document to the user's question

| User's question | Best starting point |
|---|---|
| Why might I use this package? | `README` |
| How do I complete this workflow? | Vignette |
| What does this function or argument do? | Reference documentation |

::: {.callout-tip}
Good packages provide routes for both new and returning users.
:::

## Writing vignettes

Write for a reader who may never have used your package before.

* Begin with the reader's goal, not the package structure
* Ask someone to follow the vignette without your help

For `pokemonvgc`, the Week 7 task becomes a guide: **Rescue a Pokémon Elo plot**.

::: {.callout-tip}
If a workflow is difficult to explain, the package may also be difficult to use.
:::

## Vignettes: How to

```{r}
#| eval: false

usethis::use_vignette("rescue-an-elo-plot.qmd")

```

This creates a Quarto vignette in `vignettes/` and adds the required package metadata.

# From package to audience

## Check the whole package

Documentation that renders is not enough: check the package as a complete product.

```{r}
#| eval: false

devtools::document()
devtools::check()
```

::: {.callout-important}
Fix every error and warning. Read every note and decide whether it needs action.
:::

## Make the package installable

Once the package is in a public GitHub repository, another user can install it with:

```{r}
#| eval: false

remotes::install_github("MikeLydeamore/pokemonvgc")
```

They should not need your source files, working directory, or instructions sent separately.

::: {.callout-tip}
Put a copyable installation command in the `README`.
:::

## Publish the documentation with `pkgdown`

`pkgdown` combines your package metadata and documentation into a website.

```{r}
#| eval: false

usethis::use_pkgdown_github_pages()
```

This configures a GitHub Action that rebuilds and publishes the site when the package changes.

## One package, several routes for users

::: {.center}

`DESCRIPTION` + `README` + `man/` + `vignettes/`

↓

**checked package** + **installable GitHub repository** + **pkgdown website**

:::

::: {.callout-tip}
Keep the source documents in the package; regenerate the outputs when the source changes.
:::

## Which document would you open?

1. You have never heard of the package and want to know why it exists.
2. You need to check what the `min_events` argument controls.
3. You want to reproduce the multi-step Elo plot from last week.
4. You want to install the package from GitHub.

. . .

**README → `?recent_players` → Elo-plot vignette → README installation instructions**

## Week 8 lesson

::: callout-important

## Summary

::: nonincremental

- Packages communicate through code, metadata, reference pages, examples, and longer guides.
- Write from the user's task: explain purpose, inputs, outputs, restrictions, and interpretation.
- Run `document()` and `check()` before sharing the package.
- GitHub makes the package installable; `pkgdown` makes its documentation accessible.

:::

:::

. . .

::: callout-tip

## Resources

::: nonincremental

- [R packages: the whole game](https://r-pkgs.org/whole-game.html)
- [`roxygen2` documentation](https://roxygen2.r-lib.org/)
- [`usethis` package-development helpers](https://usethis.r-lib.org/reference/index.html)
- [`pkgdown` setup](https://usethis.r-lib.org/reference/use_pkgdown.html)

:::

:::

# Appendix: Custom print methods

## Extension: Custom print method (not assessable)

What happens when you type `print(object)` in R?

::: {.fragment}
* `print` is a function

```{r}
print
```
:::

## Methods and OOP in R

`print()` is an _S3 generic_.

We can't cover S3 precisely here (come back in Advanced R Programming for that), but we can think of it as:

[a different `print` function is called depending on what `object` looks like.]{.center}

::: {.fragment}
Which method is chosen is based on the _class_ of `object`.
:::


## Methods and OOP in R

```{r}
fit <- lm(bill_length_mm ~ bill_depth_mm, data = palmerpenguins::penguins)
class(fit)
```

::: {.fragment}
```{r}
head(methods("print"))
```
:::

## Methods and OOP in R

We can prepend our own class while preserving the object's existing classes.

```{r}
class(fit) <- c("myclass", class(fit))

print.myclass <- function(x, ...) {
  cat("A fitted model with", length(coef(x)), "coefficients\n")
  invisible(x)
}

print(fit)
```

## Custom print methods

The print method selects what a user needs to see; it does not change the underlying object.

`print` methods _are_ communication too. Compare the user-facing summary with the internal component names:

```{r}
#| echo: false

fit <- lm(bill_length_mm ~ bill_depth_mm, data = palmerpenguins::penguins)
```

::: {.fragment}
```{r}
print(fit)
```

:::

::: {.fragment}
```{r}
head(names(unclass(fit)), 6)
```
:::

## Writing print methods

All of our communication principles apply to `print` methods:

* Important information first
* Unnecessary information hidden
* Conciseness
* Clarity

Often this will need multiple rounds of feedback (like all writing).

## Including print methods in packages

To include a custom print method in a package, document it with `@export`. `roxygen2` will usually generate the required S3 registration in `NAMESPACE`.

This is out of scope for the unit, but you can read more in the [`roxygen2` S3 documentation](https://roxygen2.r-lib.org/articles/rd-s3.html).
