---
pagetitle: "ETC5523: Communicating with Data"
subtitle: "From model outputs to communicative tables"
author: "Michael Lydeamore"
email: "michael.lydeamore@monash.edu"
date: "Week 5"
department: "Department of Econometrics and Business Statistics"
unit-url: "cwd.numbat.space"
footer: "ETC5523 Week 5A"
format: 
  revealjs:
    logo: images/monash-one-line-black-rgb.png
    slide-number: c
    multiplex: false
    theme: ../assets/monash.scss
    show-slide-number: all
    show-notes: true
    controls: true
    width: 1280
    height: 720
    css: [../assets/tachyons-addon.css, ../assets/custom.css]
    include-after-body: "../assets/after-body.html"
execute:
  echo: true
---

```{r, include = FALSE}
current_file <- knitr::current_input()
basename <- gsub(".[Rq]md$", "", current_file)
library(knitr)
library(kableExtra)
library(janitor)
library(broom)
library(broom.helpers)
library(tidyverse)
knitr::opts_chunk$set(
  fig.path = sprintf("images/%s/", basename),
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 3,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  cache.path = "cache/"
)
```

## <br>[`r rmarkdown::metadata$pagetitle`]{.monash-blue} {#etc5523-title background-image="images/bg-01.png"}

### `r rmarkdown::metadata$subtitle`

Lecturer: *`r rmarkdown::metadata$author`*

`r rmarkdown::metadata$department`

::: tl
<br>

<ul class="fa-ul">

<li>

[<i class="fas fa-envelope"></i>]{.fa-li}`r rmarkdown::metadata$email`

</li>

<li>

[<i class="fas fa-calendar-alt"></i>]{.fa-li} `r rmarkdown::metadata$date`

</li>

<li>

[<i class="fa-solid fa-globe"></i>]{.fa-li}<a href="`r rmarkdown::metadata[["unit-url"]]`">`r rmarkdown::metadata[["unit-url"]]`</a>

</li>

</ul>

<br>
:::

## {#aim}

::: {.callout-important }

## Aim

* Extract information from model objects  
* Select and present model results for a particular audience

:::

. . . 

::: {.callout-tip }

## Why

* Model objects contain much more information than an audience needs
* Good communication depends on selecting, interpreting and presenting the right results

:::






## 📈 Statistical models

::: incremental
* All models are approximations of an unknown data-generating process
* Their usefulness depends on the question, the data, the assumptions and the model choice

:::

::: {.fragment}

::: {.columns}

::: {.column width='50%'}

```{r, echo = FALSE, fig.height = 4, fig.width = 6}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) + 
  geom_point() + 
  labs(
    y = "Miles per gallon\n(mpg)", x = "Weight\n(wt)", 
    title = "Motor Trend Car Road Tests")
```

:::

::: {.column width='50%'}

🎯  Characterise `mpg` in terms of `wt`.

:::

:::

:::

## Fitting linear models in R

We fit the model: 

$$\texttt{mpg}_i = \beta_0 + \beta_1\texttt{wt}_i + e_i$$

::: {.columns}
::: {.column width='50%'}

::: {.fragment}

In R we fit this as 
```{r}
#| code-line-numbers: false
fit <- lm(mpg ~ wt, data = mtcars)
```

:::
:::
::: {.column width='50%'}

::: fragment

$\hat{\beta}_0 = `r round(coef(fit)[1], 3)`$ and $\hat{\beta}_1 = `r round(coef(fit)[2], 3)`$

In `mtcars`, `wt` is measured in 1,000 lb. A car that is 1,000 lb heavier is estimated to achieve **`r abs(round(coef(fit)[["wt"]], 2))` fewer mpg**, on average.

:::
:::
:::


## ℹ️ Extracting information from the fitted model

::: incremental

* A fitted model contains several kinds of information, including:
   * the model parameter estimates,
   * model-related summary statistics, e.g. $R^2$, AIC and BIC,
   * model-related values, e.g. residuals, fitted values and predictions.

* So how do you extract these values from the `fit`?
* What does `fit` even contain?

:::


## ℹ️ Extracting information from the fitted model 

```{r}
#| code-line-numbers: false
str(fit)
```


## ℹ️ Extracting information from the fitted model


Use accessor functions to retrieve model results:
```{r}
#| code-line-numbers: false
coef(fit)
```
This gives us the estimates of $\beta_0$ and $\beta_1$. 

. . . 

An estimate is more useful when accompanied by its uncertainty:

. . . 

```{r}
#| code-line-numbers: false
confint(fit)
```

## ℹ️ Extracting information from the fitted model 


You can also get a summary of the model object:

```{r}
#| code-line-numbers: false
summary(fit)
```

. . . 

The printed summary is useful for inspection, but it is not an audience-ready result.

## <i class="fas fa-broom"></i> Model objects to tidy data {.scrollable}


```{r}
#| code-line-numbers: false
broom::tidy(fit)
```
. . . 
```{r}
#| code-line-numbers: false
broom::glance(fit)
```
. . . 
```{r}
#| code-line-numbers: false
broom::augment(fit)
```

## Each `broom` verb answers a different question

| Verb | Unit of output | Use it to find... |
|---|---|---|
| `tidy(fit)` | one row per model term | estimates, standard errors and uncertainty |
| `glance(fit)` | one row per model | model-level summaries such as $R^2$, AIC and sample size |
| `augment(fit)` | one row per observation | fitted values, residuals and influence measures |

. . .

The same interface works across many model classes. `broom` selects the appropriate method from the class of the fitted object.


# <i class="fas fa-table"></i> From model object to communicative table

## Extract the model terms you need

```{r}
#| code-line-numbers: false
broom::tidy(fit, conf.int = TRUE)
```

. . .

The tidy output is data. We can now select, rename and present the results without copying values from the printed model summary.

## A default table is a starting point

```{r, echo = TRUE, results = "asis", eval = FALSE}
default_model_table <- gtsummary::tbl_regression(
  fit,
  intercept = TRUE
)

default_model_table
```

## The default table exposes the model output

```{r, echo = FALSE, results = "asis"}
default_model_table <- gtsummary::tbl_regression(
  fit,
  intercept = TRUE
)

default_model_table
```

## Critique the table for a general audience

Imagine this table will go to a technically competent, time-poor manager.

::: incremental

1. Which row answers the substantive question?
2. Can the labels be understood without seeing the R code?
3. Does the displayed precision match the strength of the evidence?
4. Is the $p$-value necessary for this audience?
5. What interpretation or limitation must accompany the table?

:::

## Refine the table for its communication purpose

```{r, echo = TRUE, results = "asis", eval = FALSE}
model_table <- gtsummary::tbl_regression(
  fit,
  intercept = FALSE,
  label = list(wt ~ "Weight (1,000 lb)"),
  estimate_fun = gtsummary::label_style_sigfig(digits = 2)
) |>
  gtsummary::modify_column_hide(columns = p.value) |>
  gtsummary::bold_labels()

model_table
```

## The revised table prioritises the useful result

```{r, echo = FALSE, results = "asis"}
model_table <- gtsummary::tbl_regression(
  fit,
  intercept = FALSE,
  label = list(wt ~ "Weight (1,000 lb)"),
  estimate_fun = gtsummary::label_style_sigfig(digits = 2)
) |>
  gtsummary::modify_column_hide(columns = p.value) |>
  gtsummary::bold_labels()

model_table
```

## Interpret the result before formatting it

::: {.callout-important}

Each additional 1,000 lb of vehicle weight is associated with **`r abs(round(coef(fit)[["wt"]], 2))` fewer mpg**, on average (95% confidence interval: `r paste(abs(round(rev(confint(fit)["wt", ]), 2)), collapse = " to ")` fewer mpg).

:::

::: incremental

* The unit of comparison is explicit.
* The interval communicates uncertainty around the estimate.
* “Associated with” does not claim that greater weight caused the difference.
* The intercept describes a hypothetical car weighing zero and is not substantively useful here.

:::

## What results should you present?

Start with the communication question, not with everything the model produced.

::: incremental

1. Which estimate or comparison answers the audience's question?
2. What uncertainty should accompany it?
3. What units, reference categories and sample size are needed?
4. Which model summary or diagnostic matters for this purpose?
5. What limitation prevents overinterpretation?

:::

## Tables and plots support different tasks

::: incremental

* Use a **table** when readers need exact values or must look up individual results.
* Use a **plot** when readers need to compare magnitudes, patterns or uncertainty quickly.
* Use both when the tasks are genuinely different—not simply to repeat the same information.

:::

# <i class="fas fa-table"></i> Some common tables

## Use the same workflow for descriptive summaries

```{r, echo = TRUE, results = "asis", eval = FALSE}
summary_table <- mtcars |>
  mutate(cyl = factor(cyl)) |>
  select(mpg, wt, cyl) |>
  gtsummary::tbl_summary(
    by = cyl,
    label = list(
      mpg ~ "Fuel economy (mpg)",
      wt ~ "Weight (1,000 lb)"
    ),
    statistic = gtsummary::all_continuous() ~ "{mean} ({sd})",
    digits = gtsummary::all_continuous() ~ 1,
    missing = "no"
  ) |>
  gtsummary::bold_labels()

summary_table
```

## A descriptive table should make comparison easy

```{r, echo = FALSE, results = "asis"}
summary_table <- mtcars |>
  mutate(cyl = factor(cyl)) |>
  select(mpg, wt, cyl) |>
  gtsummary::tbl_summary(
    by = cyl,
    label = list(
      mpg ~ "Fuel economy (mpg)",
      wt ~ "Weight (1,000 lb)"
    ),
    statistic = gtsummary::all_continuous() ~ "{mean} ({sd})",
    digits = gtsummary::all_continuous() ~ 1,
    missing = "no"
  ) |>
  gtsummary::bold_labels()

summary_table
```

::: incremental

* The columns create the comparison between cylinder groups.
* The labels state the measures and units.
* Mean (SD) is a choice—not an automatic default. Another question or distribution may call for median (IQR).

:::

## <i class="fas fa-table"></i> Descriptive summary statistics 

::: incremental

* The main goal is to give a numerical summary to give a "feel" of what the data contains. These generally should convey:
  * variables in the data,
  * the number of observations for each variable, 
  * missing values (if any),
  * the distribution, e.g. in the form of five-number summaries or counts/percentages for categorical variables. 
  
:::

## <i class="fas fa-table"></i> Descriptive summary statistics 

* For numerical variables, you may have a **correlation table** which displays the correlation coefficients of every pair of variables. Because it is symmetrical, you can omit the upper triangle.

```{r correlation-table, echo = FALSE}
options(knitr.kable.NA = '')
mtcars |>
  select(mpg, wt, hp) |>
  pivot_longer(everything()) |>
  mutate(name = factor(name, levels = c("mpg", "wt", "hp"))) |>
  group_by(name) |>
  summarise(
    Mean = mean(value),
    SD = sd(value)
  ) |>
  mutate(
    `1.` = case_when(
      name == "wt" ~ cor(mtcars$mpg, mtcars$wt),
      name == "hp" ~ cor(mtcars$mpg, mtcars$hp),
      TRUE ~ NA_real_
    ),
    `2.` = case_when(
      name == "hp" ~ cor(mtcars$wt, mtcars$hp),
      TRUE ~ NA_real_
    )
  ) |>
  knitr::kable(table.attr = "class='cor-table'") |> 
  kable_classic(full_width = FALSE)
```

## <i class="fas fa-table"></i> Descriptive summary statistics 

```{r correlation-table-code, echo=TRUE, eval=FALSE}
#| code-line-numbers: "1|2-10|11-16|17-21|22-23"
options(knitr.kable.NA = '')
mtcars |>
  select(mpg, wt, hp) |>
  pivot_longer(everything()) |>
  mutate(name = factor(name, levels = c("mpg", "wt", "hp"))) |>
  group_by(name) |>
  summarise(
    Mean = mean(value),
    SD = sd(value)
  ) |>
  mutate(
    `1.` = case_when(
      name == "wt" ~ cor(mtcars$mpg, mtcars$wt),
      name == "hp" ~ cor(mtcars$mpg, mtcars$hp),
      TRUE ~ NA_real_
    ),
    `2.` = case_when(
      name == "hp" ~ cor(mtcars$wt, mtcars$hp),
      TRUE ~ NA_real_
    )
  ) |>
  knitr::kable(table.attr = "class='cor-table'") |> 
  kable_classic(full_width = FALSE)
```

## <i class="fas fa-table"></i> Descriptive summary statistics

* You may have **cross-tabulations** (also called **contingency tables**)
```{r xtab, echo = FALSE}
mtcars |> 
  tabyl(cyl, gear) |> 
  adorn_totals(where = c("row", "col")) |> 
  adorn_percentages("all") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns() |> 
  adorn_title("combined") |> 
  knitr::kable(table.attr = "class='xtab'") |> 
  kable_styling() |> 
  row_spec(4, extra_css = "border-top: 1px solid black;") |> 
  kable_classic(full_width = FALSE)
```

## <i class="fas fa-table"></i> Descriptive summary statistics

* You may have **cross-tabulations** (also called **contingency tables**)
```{r xtab-code, eval=FALSE, echo=TRUE}
#| code-line-numbers: "2|4-5"
library(janitor)
mtcars |> 
  tabyl(cyl, gear) |> 
  adorn_totals(where = c("row", "col")) |> 
  adorn_percentages("all") |>
  adorn_pct_formatting(digits = 1) |>
  adorn_ns() |> 
  adorn_title("combined") |> 
  knitr::kable(table.attr = "class='xtab'") |> 
  kable_styling() |> 
  row_spec(4, extra_css = "border-top: 1px solid black;") |> 
  kable_classic(full_width = FALSE)
```


## <i class="fas fa-table"></i> Model-results tables

::: incremental

* Include the estimates or comparisons that answer the audience's question.
* Accompany estimates with a standard error, confidence interval or credible interval.
* State units, reference categories and sample size where they are needed for interpretation.
* Include model-level summaries such as $R^2$, AIC or BIC only when they serve the communication purpose.
* A $p$-value or significance stars do **not** replace effect size and uncertainty.

:::

# Communication considerations for tables


## <i class="fas fa-syringe" style="color:#FFC0CB"></i> Numerical precision 

* Select an appropriate precision for your goal and audience:

```{r, echo = FALSE}
set.seed(1)
mtcars |> 
  rownames_to_column() |> 
  # add some random decimal places
  mutate(wt1 = round(wt + rnorm(n())/ 1000, 5),
         wt2 = round(wt, 2),
         wt3 = round(wt, 0)) |>  
  select(rowname, wt1:wt3) |> 
  head(4) |> 
  knitr::kable(table.attr="class='xtab'",
        col.names = c("Car brand", "5 d.p.", "2 d.p.", "0 d.p.")) |> 
  kable_styling() |> 
  column_spec(2:4, width = "150px") |> 
add_header_above(c(" " = 1, "Weight" = 3)) |> 
  kable_classic(full_width = FALSE)
```



## <i class="fas fa-syringe" style="color:#FFC0CB"></i> Numerical precision

* Display trailing zeroes to match selected precision of the column:

```{r, echo = FALSE}
tibble(zero = c(0.233, 0.32, 0.4, 0.343)) |> 
  mutate(nozero = as.character(zero)) |> 
  knitr::kable(table.attr="class='xtab'", 
               align = "rr",
               col.names = c("Yes", "No")) |> 
  kable_styling() |> 
  column_spec(1:2, width = "100px") |> 
  add_header_above(c("Trailing zeroes" = 2),
                   extra_css = "border-bottom: 1px solid black;") |> 
  kable_classic(full_width = FALSE)
```


## <i class="fas fa-syringe" style="color:#FFC0CB"></i> Numerical precision 

* Change and display units as appropriate:

::: f3

```{r, echo = FALSE}
mtcars |> 
  rownames_to_column() |> 
  # add some random decimal places
  mutate(wt1 = wt * 1000,
         wt2 = wt,
         wt3 = round(wt1 * 0.453592),
         wt4 = signif(wt3 * 1000, 3),
         wt5 = scales::comma(signif(wt3 * 1000000))) |>  
  select(rowname, wt1:wt5) |> 
  head(4) |> 
  knitr::kable(table.attr="class='xtab'",
               col.names = c("Car brand", "Weight (lb)", "Weight (1,000 lb)", "Weight (kg)", "Weight (g)", "Weight (mg)"),
               align = "lrrrrr") |> 
  kable_styling() |> 
  column_spec(2:6, width = "200px") |> 
  kable_classic()
  
```

:::

* Use digit-grouping conventions familiar to your audience. <br>For example, `1,000,000` is easier to read than `1000000` for an Australian audience.

```{r}
scales::comma(c(439024, 4900343), accuracy = 1000)
```


## <i class="fas fa-align-left"></i> Column <i class="fas fa-align-center"></i> alignment   <i class="fas fa-align-right"></i>

* Centre spanner labels.
* Right-align numbers.
* Left-align text.

::: f3

```{r, echo = FALSE}
mtcars |> 
  rownames_to_column("car") |> 
  # add some random decimal places
  mutate(
    car1 = car,
    car2 = car,
    car3 = car,
    hp1 = hp,
    hp2 = hp,
    hp3 = hp) |>  
  select(car1:hp3) |> 
  head(4) |> 
  knitr::kable(table.attr="class='xtab'",
               col.names = c("Left", "Center", "Right", "Left", "Center", "Right"),
               align = "lcrlcr") |> 
  kable_styling() |> 
  column_spec(1:3, width = "200px") |> 
  column_spec(4:6, width = "80px") |> 
  add_header_above(c("Car brand" = 3,
                     "Horsepower" = 3)) |> 
  kable_classic()
```

:::


## <i class="fas fa-align-justify"></i> Labels within tables

* Tables intended as final products should use clear, descriptive labels.
* Put the unit once in the column heading.
* Do not repeat the unit in every cell.

::: {.f3}

<center>
❌
</center>

```{r, echo = FALSE}
mtcars |> 
  rownames_to_column("car") |> 
  select(car, wt, disp) |> 
  mutate(wt = paste(wt*1000, "lb"),
         disp = paste(disp, "cubic inches")) |> 
  head() |> 
  knitr::kable() |> 
  kable_classic(full_width = FALSE)
```
:::


## <i class="fas fa-align-justify"></i> Labels within tables

* Tables intended as final products should use clear, descriptive labels.
* Put the unit once in the column heading.
* Do not repeat the unit in every cell.

::: {.f3}

<center>
✅
</center>


```{r, echo = FALSE}
mtcars |> 
  rownames_to_column("car") |> 
  select(car, wt, disp) |> 
  head() |> 
  knitr::kable(col.names = c("Car brand", "Weight (1,000 lb)", "Displacement (cubic inches)"), escape = FALSE) |>
  kable_classic(full_width = FALSE)
```

:::





## Week 5 Lesson 


::: callout-important

## Summary

* Use accessors and `broom` to turn a fitted model into data you can work with.
* Select results according to the audience's question—not simply because the model produced them.
* Communicate estimates with uncertainty, units, context and appropriate limitations.
* Use `gtsummary` to create a table, then refine its precision, alignment and labels.

:::


::: callout-tip

## Resources

* [`broom` documentation](https://broom.tidymodels.org/)
* [`gtsummary` for summary tables](https://www.danieldsjoberg.com/gtsummary/)

:::
